### 2022-06-28 code/old/news_framing.Rmd

---
title: "Factor analysis migration framing"
author: "Nicolai Berk"
date: "7 7 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = T)

# import libraries
library(dplyr)
library(tidyr)
library(stringr)
library(tidytext)
library(data.table)
library(here)
library(stm)

# library(psych)
# library(GPArotation)
# library(psy)
# library(corrplot)
# library(anacor)

# load data
mig_dta <- fread(here("data/_migration_btgrams.csv")) # migration articles with bi- and trigram dtm

# clean
mig_dta$paper[mig_dta$paper == "weltonline"] <- "welt"
mig_dta$date_new <- NA
mig_dta$date_new[mig_dta$paper == "bild"] <- mig_dta$date[mig_dta$paper == "bild"] %>% as.Date()
mig_dta$date_new[mig_dta$paper == "faz"] <- mig_dta$date[mig_dta$paper=="faz"] %>% as.integer() %>% lubridate::days() + as.Date("1970-01-01")
mig_dta$date_new[mig_dta$paper == "spon"] <- mig_dta$date[mig_dta$paper=="spon"] %>% as.Date(format = "%d.%m.%Y")
mig_dta$date_new[mig_dta$paper == "sz"] <- mig_dta$date[mig_dta$paper=="sz"] %>% as.integer() %>% lubridate::days() + as.Date("1970-01-01")
mig_dta$date_new[mig_dta$paper == "taz"] <- mig_dta$date[mig_dta$paper=="taz"] %>% as.integer() %>% lubridate::days() + as.Date("1970-01-01")
mig_dta$date_new[mig_dta$paper == "welt"] <- mig_dta$date[mig_dta$paper=="welt"] %>% as.integer() %>% lubridate::days() + as.Date("1970-01-01")
mig_dta$date_new <- as.Date(mig_dta$date_new, origin = '1970-01-01')
mig_dta <- mig_dta[mig_dta$date_new > as.Date('2010-01-01'),]

```

<!-- ## Factor analysis -->

<!-- ```{r preprocessing} -->
<!-- items <- mig_dta[complete.cases(mig_dta[, 9:(ncol(mig_dta)-1)]), 9:(ncol(mig_dta)-1)] -->

<!-- # smoothing -->
<!-- items <- items + .01 -->
<!-- items <- log(items/sum(items)) -->

<!-- cormtrx <- cor(items) -->

<!-- ``` -->

<!-- Calculate eigenvalues -->
<!-- ```{r eigen} -->
<!-- # eigen_d <- eigen(cormtrx) -->
<!-- # sum(eigen_d$values > 1.5) -->
<!-- # nfactors(items) -->
<!-- ``` -->

<!-- Scree plot -->
<!-- ```{r scree} -->
<!-- scree.plot(eigen_d$values, type = "E") -->
<!-- scree.plot(eigen_d$values[1:100], type="E") -->
<!-- scree.plot(eigen_d$values[1:50], type="E")  -->

<!-- ``` -->

<!-- No clear elbow, far too many factors with eigenvalue $>$ 1.5, trying with 8 factors for now based on scree plot. -->

<!-- ```{r fa_model} -->
<!-- # fa_model <- fa(cormtrx, n.obs = nrow(items), nfactors=8, use = "pairwise.complete.obs") -->
<!-- fa_model <- factanal(items, 8, rotation="varimax") -->

<!-- ``` -->

<!-- Are the factors meaningful? - assessment of 20 items with highest correlation for each factor -->

<!-- ```{r fa_asses} -->

<!-- ``` -->

<!-- ## Correspondence Analysis -->

<!-- Since factor analysis doesn't seem to work at the moment, we'll try to make do with correspondence analysis, as suggested by @Lowe2016. -->

<!-- [Try to figure out how to plot sigma (canonical correlation), see @Lowe2016, Fig. 3] -->

<!-- ```{r ca} -->
<!-- res <- anacor(cormtrx) -->
<!-- summary(res) -->

<!-- ``` -->

## STM

As I am interested in different topics within the discourse about immigration rather than the different positions on immigration, I opt for the commonly used STM (Stewart 2016).

```{r stm}

## assess token count
token_count <- mig_dta[,9:(ncol(mig_dta)-1)] %>% 
  colSums()
names(token_count) <- colnames(items)
token_count %>% sort(decreasing = T) %>% head(30)

## pre-processing
mig_dta <- mig_dta[complete.cases(mig_dta),]
mig_dta <- mig_dta[rowSums(mig_dta[, 9:(ncol(mig_dta)-1)]) > 0, ] # drop empty documents
meta <- mig_dta[, c('date_new', 'paper')]
mig_dtm <- quanteda::as.dfm(mig_dta[, 9:(ncol(mig_dta)-1)])

stm_model <- stm(documents = mig_dtm,
                 K = 0, # sets an ideal number of topics, however should be tested with searchK()
                 prevalence =~ paper * date_new, 
                 max.em.its = 75, 
                 data = meta, 
                 init.type = "Spectral")

labelTopics(stm_model)

```

## Migration Framing of Bild vs others

## AfD slant in migration coverage