### 2022-06-28 code/old/robust.Rmd

---
title: "Robustness and additional estimates"
author: "Nicolai Berk"
date: "14 3 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_chunk$set(echo = FALSE, message = F, warning = F, cache = T, fig.align='center', eval = T, dev = "svg")
# knitr::opts_knit$set(root.dir = 'C:/Users/nicol/Dropbox/PhD/Papers/Bild/github/bild')

library(tidyverse)
library(dplyr)
library(data.table)
library(fixest)
library(lubridate)
library(patchwork)
```

## Why is the direct effect of the crime treatment negative? Shouldn't it be positive?

### Standard model

```{r}
gles_p_long <- fread('C:/Users/nicol/Dropbox/PhD/Papers/Bild/github/bild/data/gles/Panel/long_cleaned.csv')

gles_p_long$dv <- gles_p_long$`1130_clean`
gles_p_long$treat <- gles_p_long$`1661a_clean` > 0
gles_p_long$treat[is.na(gles_p_long$`1661a_clean`)] <- NA

gles_p_long$post <- gles_p_long$date_clean>= as.Date("2017-02-01")
gles_p_long <- 
  gles_p_long %>% 
  select(lfdn, date_clean, `1130_clean`) %>% 
  filter(!is.na(`1130_clean`)) %>% 
  arrange(lfdn, date_clean) %>% 
  group_by(lfdn) %>% 
  slice(1) %>% 
  mutate(init_mig = `1130_clean`) %>% 
  select(lfdn, init_mig) %>% 
  right_join(gles_p_long, by = "lfdn")

gles_p_long <- 
  gles_p_long %>% 
  mutate(mip_mig_crime_lag = lag(mip_mig_crime)) %>% 
  filter(!is.na(mip_mig_crime)) %>% 
  arrange(lfdn, date_clean) %>% 
  group_by(lfdn) %>% 
  slice(1) %>% 
  mutate(mip_mig_crime_init = mip_mig_crime) %>% 
  select(lfdn, mip_mig_crime_init) %>% 
  right_join(gles_p_long, by = "lfdn")

gles_p_long_1617 <- 
  gles_p_long %>% 
        filter(date_clean >= as.Date("2016-01-01"),
               date_clean <  as.Date("2018-01-01"))
rm(gles_p_long)

# model 1: direct effect
ate <- 
  feglm(dv ~ post * treat | lfdn, 
      cluster = c("lfdn"), 
      data = gles_p_long_1617)

inter <- 
  feglm(dv ~ post * treat * init_mig, 
      cluster = c("lfdn"), 
      data = gles_p_long_1617)

points <- ate$coefficients
ests <- cbind(points, confint(ate))
plot_did_ate <- 
  ests %>% 
  # mutate(name = factor(row.names(.), levels = row.names(.))) %>% 
  ggplot(aes(y = row.names(.), x = points, xmin = `2.5 %`, xmax = `97.5 %`)) +
  geom_pointrange() +
  geom_vline(xintercept = 0, col = "red", lty = 2) +
  ylab("") + xlab("Estimate") +
  scale_y_discrete(limits=rev(row.names(ests)),
                   labels = rev(c("Post", "Treatment", "Post X Treatment")))

points <- inter$coefficients
ests <- cbind(points, confint(inter))

plot_did_inter <- 
  ests %>% 
  filter(row.names(.) %in% c("postTRUE:treatTRUE", "postTRUE:treatTRUE:init_mig")) %>% 
  mutate(name = fct_reorder(row.names(.), desc(row.names(.)))) %>% 
  ggplot(aes(y = name, x = points, xmin = `2.5 %`, xmax = `97.5 %`)) +
  geom_pointrange() +
  geom_vline(xintercept = 0, col = "red", lty = 2) +
  ylab("") + xlab("Estimate") +
  ggtitle("DiD-estimate with interaction") +
  xlim(-0.1, 0.4) + 
  scale_y_discrete(labels = rev(c("Post X Treatment", "Post X Treatment X Initial Opinion")))

plot_did_ate/plot_did_inter

```

## Direct effect, controlling for initial attitude

```{r}

ctrl <- 
  feglm(dv ~ post * treat + init_mig, 
      cluster = c("lfdn"), 
      data = gles_p_long_1617)

points <- ctrl$coefficients
ests <- cbind(points, confint(ctrl))
plot_did_ctrl <- 
  ests %>% 
  mutate(name = factor(row.names(.), levels = row.names(.))) %>%
  filter(!name %in% c("", "(Intercept)", "init_mig")) %>% 
  ggplot(aes(y = row.names(.), x = points, xmin = `2.5 %`, xmax = `97.5 %`)) +
  geom_pointrange() +
  geom_vline(xintercept = 0, col = "red", lty = 2) +
  ylab("") + xlab("Estimate") +
  scale_y_discrete(limits=rev(row.names(ests)[!row.names(ests) %in% c("", "(Intercept)", "init_mig")]),
                   labels = rev(c("Post", "Treatment", "Post X Treatment"))) +
  xlim(-0.1, 0.4) + 
  ggtitle("Controlling for initial attitude")

plot_did_ctrl

```

## Distribution of Bild readers on pre-treatment outcome

```{r}

bild_readers <- 
  gles_p_long_1617 %>% 
  arrange(lfdn, date_clean) %>% 
  group_by(lfdn) %>% 
  slice(1) %>% 
  filter(`1661a_bin`) %>% 
  ggplot(aes(x = `1130_clean`, after_stat(density))) +
  geom_histogram(binwidth = .5) + 
  ggtitle("Bild readers") + xlab("") +
  theme_minimal() +
  ylim(0, 0.75)

rest_readers <- 
  gles_p_long_1617 %>% 
  arrange(lfdn, date_clean) %>% 
  group_by(lfdn) %>% 
  slice(1) %>% 
  filter(!`1661a_bin`) %>% 
  ggplot(aes(x = `1130_clean`, after_stat(density))) +
  geom_histogram(binwidth = .5) + 
  ggtitle("Not bild readers") +xlab("") +
  theme_minimal() +
  ylim(0, 0.75)

bild_readers/rest_readers

```

## estimate effect based on pre-treatment groups

```{r}

gles_p_long_1617 <- 
  gles_p_long_1617 %>% 
  filter(wave == "1") %>% 
  mutate(bild_init = ifelse(is.na(`1661a_clean`), NA, `1661a_clean` > 0)) %>% 
  select(bild_init, lfdn) %>% 
  right_join(gles_p_long_1617, by = "lfdn")

treat_alt_ate <- 
  feglm(dv ~ post * bild_init, 
      cluster = c("lfdn"), 
      data = gles_p_long_1617)

points <- treat_alt_ate$coefficients
ests <- cbind(points, confint(treat_alt_ate))
plot_did_alt_ate <- 
  ests %>% 
  mutate(name = factor(row.names(.), levels = row.names(.))) %>%
  filter(!name %in% c("", "(Intercept)", "init_mig")) %>% 
  ggplot(aes(y = row.names(.), x = points, xmin = `2.5 %`, xmax = `97.5 %`)) +
  geom_pointrange() +
  geom_vline(xintercept = 0, col = "red", lty = 2) +
  ylab("") + xlab("Estimate") +
  scale_y_discrete(limits=rev(row.names(ests)[!row.names(ests) %in% c("", "(Intercept)", "init_mig")]),
                   labels = rev(c("Post", "Treatment", "Post X Treatment"))) +
  xlim(-0.1, 0.4) + 
  ggtitle("Controlling for initial attitude")

plot_did_alt_ate

```

## interaction with alternative treatment definition

```{r}

treat_alt_inter <- 
  feglm(dv ~ post * bild_init * init_mig, 
      cluster = c("lfdn"), 
      data = gles_p_long_1617)

points <- treat_alt_inter$coefficients
ests <- cbind(points, confint(treat_alt_inter))
plot_did_alt_inter <- 
  ests %>% 
  mutate(name = factor(row.names(.), levels = row.names(.))) %>%
  filter(name %in% c("postTRUE:bild_initTRUE", "postTRUE:bild_initTRUE:init_mig")) %>% 
  ggplot(aes(y = row.names(.), x = points, xmin = `2.5 %`, xmax = `97.5 %`)) +
  geom_pointrange() +
  geom_vline(xintercept = 0, col = "red", lty = 2) +
  ylab("") + xlab("Estimate") +
  scale_y_discrete(limits=rev(row.names(ests)[row.names(ests) %in% c("postTRUE:bild_initTRUE", "postTRUE:bild_initTRUE:init_mig")]),
                   labels = rev(c("Post X Treatment", "Post X Treatment X Initial Attitude"))) +
  xlim(-0.1, 0.4) + 
  ggtitle("Controlling for initial attitude")

plot_did_alt_inter

```

## Nicer visualisation of interaction effect

### new treatment definition

```{r}

## run bootstrap to estimate 1000 direct and conditional effects

library(rsample)
library(purrr)

set.seed(42)
B <- 1000

modeldata <- 
  gles_p_long_1617 %>% 
  filter(!is.na(bild_init) & !is.na(dv) & !is.na(init_mig)) %>% 
  select(lfdn, dv, init_mig, bild_init, post) %>% 
  nest(-lfdn)

bs <- bootstraps(modeldata, times = B)
bs_ests <- map(bs$splits, ~as.tibble(.) %>% 
                       unnest %>% 
                       summarize(DiD = 
                                   feglm(dv ~ post * bild_init * init_mig, 
                                    cluster = c("lfdn"), 
                                    data = .) %>% 
                                   coef() %>% 
                                   .["postTRUE:bild_initTRUE"],
                                 IA = 
                                   feglm(dv ~ post * bild_init * init_mig, 
                                    cluster = c("lfdn"), 
                                    data = .) %>% 
                                   coef() %>% 
                                   .["postTRUE:bild_initTRUE:init_mig"],
                                 )) %>% 
  bind_rows(.id = 'boots')

direct <- coef(treat_alt_inter)["postTRUE:bild_initTRUE"]
interact <- coef(treat_alt_inter)["postTRUE:bild_initTRUE:init_mig"]

ggplot() +
  geom_histogram(data = gles_p_long_1617 %>% filter(wave == "1"), mapping = aes(x = init_mig, after_stat(density)), alpha = 0.4, bins = 8) +
  geom_abline(data = bs_ests, aes(slope = IA, intercept = DiD, group = boots), alpha= 0.05) +
  geom_abline(slope = interact, intercept = direct, col = "red") +
  theme_minimal() +
  geom_hline(yintercept = 0, col = "black") +
  scale_x_continuous(limits=c(-3,3)) +
  scale_y_continuous(limits=c(0,.6)) +
  ggtitle(paste0(B, " bootstrap estimates of heterogenous treatment effect")) +
  ylab("Estimate") + xlab("Initial migration attitude")

ggsave('C:/Users/nicol/Dropbox/PhD/Papers/Bild/github/bild/paper/vis/DiDXinit_imm.svg', device = "svg")

```

### old treatment definition (probably shouldn't be used anymore)