### 2022-06-28 pres/20211028/EPSIP_20211028.Rmd

---
title: "The effect of changing news frames"
author: "Nicolai Berk$^*$"
institute: "$^*$RTG Dynamics/Humboldt Universität zu Berlin"
date: "Oct 14th, 2021"
output: 
  beamer_presentation:
    theme: "Frankfurt"
    # colortheme: "lily"
    # fonttheme: "professionalfonts"
    slide_level: 2
    toc: False
header-includes:
# - \usepackage{fontawesome}
- \AtBeginSection{}
bibliography: PhD-Proposal.bib
---

# Motivation

\center

![](vis/volksempfanger.jpg){width=30%} 
![](vis/bild_umgebracht_asyl.jpeg){width=60%}

# Literature and Theory

## Media effects - maximal or minimal? {.allowframebreaks}

Maximal paradigm also follows from classic work:

- attitude instability [@Converse1962; @Zaller1992]
- agenda-setting [@McCombs1972]
- framing [@Nelson1997]

\framebreak

Minimal effects assumption with increasing media diversity [@Bennett2008]:

- media environment more polarised and more diverse,
- viewers more likely to reject news conflicting with their views,
- viewers can opt for other sources.

$\rightarrow$ strong effects unlikely

## Empirical evidence {.allowframebreaks}

### Evidence for large effects

- Framing effects found in experimental data [@Busby2019; @Leeper2020].
- Newspaper slant affects attitudes [@Foos2020].
- Salience and tonality of migration news affects attitudes [@Boomgaarden2009].

### Evidence for no/weak effects

- Newspaper slant has no effect on attitudes [@Gentzkow2011; @Guess2021; @Stetka2020].
- Newspaper takeovers have no effect on attitudes [@Durante2012; @Spirig2020].

## Synthesis

**When** can we expect strong media effects?

- Many current approaches test media effects of different outlets [@Gentzkow2011; @Guess2021].
- However, \alert{news consumers discount bias and take cues from outlets}.
  + @Baum2008 show that consumers take heuristics about content bias from outlet brands.
  + @Chiang2011a show that outlet bias moderates the effect of candidate endorsements.

$\rightarrow$ \alert{within-outlet changes} in content most likely cases to observe media effects.

## Emphasis framing {.allowframebreaks}

**What kind of changes should matter?**

- I draw on literature on 
  + the \alert{value-expectancy} framework [@Ajzen2000].
  + and \alert{emphasis framing} [@Leeper2020; @Nelson1997].
- Argues that issue attitudes are a product of \alert{associated considerations}.

<!-- \framebreak -->

<!-- - For example, migration could be associated with  -->
<!--   + humanitarian crises in the mediterranean (positive frame);  -->
<!--   + crime (negative frame); -->
<!--   + cost for the welfare state (negative frame); -->
<!--   + labour market demand (positive frame). -->
<!-- - These considerations will be weighted to form an overall evaluation of the issue at hand. -->

\framebreak

![](vis/CognitiveStorage.png)

# Empirical strategy

## The case: Germany 2017

- First election after 2015 refugee movements.
- Radical-right challenger party enters parliament, center loses.
- Migration top of the agenda.

$\rightarrow$ constantly strong attention to migration,

$\rightarrow$ good case to study effects of framing.

## Design

- Collect 2.5M articles from major German newspapers.
- Classify according to migration content.
- Identify emphasis frames.
- Correlate changes in framing with changes in attitudes.

## Issue attitudes and news consumption {.allowframebreaks}

- Panel data from the German Longitudinal Election Study [@GLES2019LongTermTracking].
- Contains questions on news consumption, immigration and integration attitudes, as well as party preference.
- Dependent variables:
  + Immigration attitude (7-point scale).
  + Integration attitude (7-point scale).
  + Attitude toward AfD (11-point scale).

## Migration content {.allowframebreaks}

- Pre-assign likelihoods for migration content using extended migration dictionary.
- Draw stratified sample of 1,800 articles, hand-code.
- Fine-tune German BERT deep-learning classifier.
- Performs very well: F1: 0.94, recall: 0.93, precision: 0.95.
- 13.5k out of 400k articles in 2017 about migration (3.5\%)

\framebreak

![](vis/salience_papers.png)

## Migration framing {.allowframebreaks}

- Estimate 60-topic structural topic model [@Roberts2014], using date and paper as covariates.
- Annotate.
- Select relevant frames with clear expectations regarding attitudinal effects.

\framebreak

\alert{Topics}:

- Capital/sexual crime committed by refugees,
- general crime,
- refugee numbers,
- labour market needs for and job market integration of refugees,
- deportations,
- internment camps (e.g. Moria),
- drownings in the Mediterranean.

\framebreak

![](vis/frames_papers_pres.png)

## Estimation

*Two models*:

- OLS of aggregate Difference-in-Differences (DiDs).
- Individual-level model with 2-way fixed effects.

## DiD-model {.allowframebreaks}

$$y = \beta_1*W + \beta_2*R + \beta_3*W*R$$

\centering

![](vis/Illustration_of_Difference_in_Differences.png){width=50%}

\framebreak

- Estimate change in newspaper framing and attitudes among readers from one wave to another, controlling for shifts in other newspapers/reader groups.
- Regress opinion shift among readership on shift in newspaper attention to different frames.
- Better identified (exact change *beyond* general trend), 
- but framing not individually matched.

## 2-way FE model {.allowframebreaks}

- Individual estimate of frame attention for each respondent, according to newspaper read.
- Regress opinion on exposure to each frame, 
- controlling for wave and individual fixed-effects.

\framebreak

- Individual estimates,
- but explains *all* variation beyond 
  + time-independent individual factors and 
  + general trends across time,
  + including random individual deviations.

# Preliminary results

Different specifications:

- all readers/exclusive readers of one newspaper,
- different lags to measure exposure (1 day, 1 week, 1 month, half a year),
- immigration and integration attitude as dependent variable.

## Difference-in-Difference model {.allowframebreaks}

![](vis/frame_effects_did.png)

## Fixed-effect model

![](vis/frame_effects_fe.png)

## Preliminary conclusion {.allowframebreaks}

\begin{center}

\textbf{No effects of media framing even when considering within-outlet changes.}

\end{center}

\framebreak

However, other possible explanations:

- Issue-sorting in 2015, subsequent "digging in",
- Measurement issues (esp. news exposure).

$\rightarrow$ further work necessary...

## Next steps (possibly paper #2)

### Theoretical additions:

- Test if readers respond by **stopping to read newspaper** [@Arceneaux2013; @Chiang2011a].
- Assess **issue sorting** [@Carmines1986] across time.

### Empirical additions:

- (Alternative measure of association: diachronic embeddings? embedding topic model?)
- (Back to natural experiment?)
- (More cases?)

## Fin

\centering

*Thank you!*

# Appendix

## Variation dependent variable {.allowframebreaks}

\centering

![](vis/imm_readers_pres.png)

\framebreak

\centering

![](vis/DiD_model_ps_immint.png)

## Potential natural experiment

\centering

![](vis/frames_bild_diekmann_pres.png)

## Salience {.allowframebreaks}

![](vis/salience_papers_focus.png)

## Framing attention {.allowframebreaks}

![](vis/frames_papers_focus_pres.png)

## Specifications DiD

\centering

![](vis/effectplot_frames_did.png){width=70%}

## Specifications 2-way FE

\centering

![](vis/effectplot_frames_fe_abs.png){width=70%}

# Resources

## Resources {.allowframebreaks}