### 2022-06-28 pres/20211126/PolBB_261121.Rmd

---
title: "Assessing media influence in contemporary news environments"
subtitle: "Right-wing media, crime, and anti-immigration attitudes"
author: "Nicolai Berk"
institute: "Humboldt Universität Berlin"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    seal: false
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
    includes:
      after_body: insert-logo.html
---

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#03396d", background_color = "#bfbfbf",
  base_font_size = "25px",
  header_h1_font_size = "2rem",
  header_h2_font_size = "1.75rem",
  header_h3_font_size = "1.25rem",
  header_font_google = google_font("Montserrat", 'Thin 100'),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono"),
  footnote_font_size = "0.75rem"
)
```

class: inverse, center, middle

# Media Influence in Contemporary News Environments
### Right-wing Media, Crime, and Anti-Immigration Attitudes

<br></br>

### Nicolai Berk<sup>*</sup>

Presentation at the Political Behavior Workshop Berlin-Brandenburg, WZB

`r Sys.Date()`

.left[.footnote[<sup>*</sup> Dynamics RTG & Humboldt Universität Berlin]]

---
class: middle, center, inverse

# Can the news media affect voters' political attitudes and behaviour?

---

# Can the news media shape opinions?

### Maximal effects literature builds on 
  + idea that __media is the main source of information__ for citizens,
  + experimental findings,
  + attitude instability (Converse; Zaller).

### Minimal effects scholars argue that 
  + __selective exposure__ 
  + in diverse and polarised contemporary news markets
  + make opinion changes unlikely.
  + Consumers will instead __change outlet__.

---

# Empirical literature

### Focus on

- __changing supply__ in media markets (Foos & Bischof 2020; Gentzkow, Shapiro & Silkinson 2011; Grossman, Margalit & Mitts 2021),
- __highly visible__ media takeovers (Durante & Knight 2012; Spirig 2020).

--

### Findings:

- __Consumers change outlet, but not opinion__ when takeovers are salient (Durante & Knight 2012; Spirig 2020).
- __Consumers persuaded__ when they change outlet (Grossman, Margalit & Mitts 2021).
- __Consumers socialised__ in new media environment will change opinion (Foos & Bischof 2020).

---

# Blind spots I

Highly visible takeovers & media changes such as boycotts result in __co-treatment__.

### Example: Sun Boycott in Merseyside (Foos & Bischof 2021)

> "_You’re in Liverpool mate, not sure if you’re a football fan, but no Scouser would dare read it. You might get the odd one or two in the city, maybe the odd tourist, but that’s about it. Ever since Hillsborough, nobody in this city touches it._"  (p. 5).

---
# Blind spots II

### Unclear whether

- __less visible changes__ in the framing of a media outlet might __persuade__.
- short-term effects ( __persuasion__ ) can be observed __or socialisation__ is main mechanism.
- changes in __framing or co-treatment__ (esp. identity changes) are main driver.

<!-- __First evidence on less visible changes in media framing suggests persuasive effects (Djourelova 2021).__ -->

---

class: center, inverse, middle

# Are changes in framing more persuasive when they are less visible?

---

# The Case

.left-column[

![Source: Wikipedia](https://upload.wikimedia.org/wikipedia/commons/thumb/3/3c/Kai_Diekmann_MediaMakers.jpg/435px-Kai_Diekmann_MediaMakers.jpg)

![Source: Wikipedia](https://upload.wikimedia.org/wikipedia/commons/thumb/5/5a/2018-11-07-Julian_Reichelt-Maischberger-1141.jpg/800px-2018-11-07-Julian_Reichelt-Maischberger-1141.jpg)

]

.right-column[

- Bild editor-in-chief Kai Diekmann drove an unexpectedly positive slant towards refugees in 2015.
- Diekmann left in February 2017, with Reichelt taking over.
- Reichelt increasingly **framed migrants as criminals** (Niggemeier 2018).

Interesting because:

- __Low-salience__ change in editorial slant.
- clearly identifiable short-term-change in framing.
$\rightarrow$ __No co-treatment__!

]

---
# Emphasis framing

**Why should changing definitions matter?**

- I draw on literature on 
  + the __value-expectancy framework__ (Ajzen & Fishbein 2000).
  + and __emphasis framing__ (Leeper and Slothuus 2020; Nelson et al. 1997).
- Argues that issue attitudes are a product of __associated considerations__.

---
class: center, middle

.left[
# Emphasis framing II 
]

```{r emph_framing, eval=TRUE, echo = F, out.width="70%", fig.pos="center"}

knitr::include_graphics(path = "vis/CognitiveStorage.png")

```

---
class: inverse, center, middle

# Empirical Strategy

---
# Measuring Framing

### Identify migration content

- Annotate 1800 articles to identify their migration content.
- Fine-tune German BERT deep-learning classifier.
- accuracy: 95.5% (F1: 0.94, recall: 0.93, precision: 0.95)
- 13.5k out of 400k articles in 2017 about migration (3.5\%)

### Measure migration frames

- Estimate 60-topic structural topic model (Roberts, Stewart and Tingley 2019), using date and paper as covariates.
- Annotate.
- Select relevant frames with clear expectations regarding attitudinal effects.

---
class: inverse, center, middle

```{r salience_des, eval=TRUE, echo = F, out.width="70%", fig.pos="center"}

knitr::include_graphics(path = "vis/salience_papers.png")

```

---
class: inverse, center, middle

```{r frames_des, eval=TRUE, echo = F, out.width="70%", fig.pos="center"}

knitr::include_graphics(path = "vis/frames_papers_pres.png")

```

---
# The treatment I

```{r treatment, eval=TRUE, echo = F, out.width="70%", fig.pos="center"}

knitr::include_graphics(path = "vis/frames_bild_diekmann_pres.png")

```

---
# The treatment II

```{r treatment2, eval=TRUE, echo = F, out.width="70%", fig.pos="center"}

knitr::include_graphics(path = "vis/did_papers_frames_1617_scatter.png")

```

---
# The treatment III

```{r treatment3, eval=TRUE, echo = F, out.width="70%", fig.pos="center"}

knitr::include_graphics(path = "vis/did_papers_allframes_all_hist.png")

```

---
# Estimation

Two studies:

- Study correlating changes in migration framing with readers' migration opinion in *survey data*.
  + Main advantage: direct measurement of dependent, individual level data.
  + Disadvantage: Self-reported newspaper exposure.
- DiD assessing impact of changing framing on *local-level proxies* like voting behaviour, protests and violence.
  + Main advantage: Proper causal identification (no limitation of survey data).
  + Disadvantage: Only aggregate-level measures.

---
class: inverse, center, middle

# First Results

---

# Individual-level study

Different specifications:

- DiD only for Bild shift, changing frame attention per newspaper.
- all readers/exclusive readers of one newspaper,
- different lags to measure exposure (1 day, 1 week, 1 month, half a year),
- immigration and integration attitude as dependent variable.

---
# 2-way fixed-effects model

$$y = \beta_1*ID + \beta_2*Wave + \beta_{f1}*ExposureFrame_1 + ... + \beta_{fi}*ExposureFrame_i$$

```{r effects_fe, eval=TRUE, echo = F, out.width="70%", fig.pos="center"}

knitr::include_graphics(path = "vis/frame_effects_fe.png")

```

---
# Difference-in-Difference correlation

$$y = \beta_1*W + \beta_2*R + \beta_3*W*R$$
```{r effects_did, eval=TRUE, echo = F, out.width="70%", fig.pos="center"}

knitr::include_graphics(path = "vis/frame_effects_did.png")

```

---
# simple DiD Bild shift

.left-column[

- Fun friday morning surprise!
- Effects in wrong direction.
- __Think about control group__

]

.right-column[

```{r simple_DiD, echo = F, message=F, fig.pos = 'center'}

library(dplyr)
library(ggplot2)
library(here)

did_ests <- read.csv(here('data/simple_did.csv'))

did_ests %>% 
  ggplot(aes(x = est,
             xmin = lower,
             xmax = upper,
             y = control)) +
  geom_pointrange() +
  geom_vline(xintercept = 0, col = 'red', lty = 2) +
  facet_grid(period~dv) +
  xlab('Estimated change on 7-point scale')+
  ylab('Control group')+
  ggtitle('Estimated opinion change following Reichelt takeover', 'For different specifications')

```

]

---
class: inverse, center, middle

# Next steps

---
# Estimation on the local level

$$y_{dt} = \beta_1 * Bild_{d} + \beta_2 * Post_{t} + \beta_3 * Bild_{d} * Post_{t}$$

Potential anti-immigation attitude proxies:

- Voting behaviour
- Anti-immigrant/-refugee violence
- Protests
- Google Searches

---
# Open questions

- Framing
  + Get rid of emphasis framing? 
  + More focus on crime as right-wing issue?
  + More electoral behaviour focus?
- Appropriate control group?
- Matching? Instruments?
- Add 2015 framing shift as second case?

---
class: center, middle, inverse

# Thank you!

---

<!-- # News media framing -->

<!-- .left-column[ -->
<!-- ![via http://tabloid-watch.blogspot.com/2013/01/corrections-must-be-given-more.html](vis/130103mail.jpg) -->

<!-- ![via ](vis/bild_umgebracht_asyl.jpeg) -->

<!-- .right-column[ -->

<!-- ```{r bild_main_2, eval=TRUE, echo = F, out.width="70%", fig.pos="center"} -->

<!-- knitr::include_graphics(path = "vis/bild_auslaender.jpeg") -->

<!-- ``` -->

<!-- .footnote[via https://uebermedien.de/1184/die-bruellspirale/] -->