### 2022-06-28 pres/20220128/UNDA_20220128.Rmd

---
title: "Assessing media influence in contemporary news environments"
subtitle: "Right-wing media, crime, and anti-immigration attitudes"
author: "Nicolai Berk"
institute: "Humboldt Universität Berlin"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    seal: false
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
    includes:
      after_body: insert-logo.html
bibliography: PhD-Proposal.bib
---

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#03396d", background_color = "#bfbfbf",
  base_font_size = "25px",
  header_h1_font_size = "2rem",
  header_h2_font_size = "1.75rem",
  header_h3_font_size = "1.25rem",
  header_font_google = google_font("Montserrat", 'Thin 100'),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono"),
  footnote_font_size = "0.75rem"
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = F, warning = F, dev.args = list(bg = 'transparent') )
```

```{r bib, include=F, warning = F}
library(RefManageR)
BibOptions(check.entries = FALSE, bib.style = "authoryear", style = "text")
# file <- system.file("Bib", "PhD-Proposal.bib", package = "RefManageR")
bib <- ReadBib("PhD-Proposal.bib", check = F)

```

class: inverse, center, middle

# Puppet Masters, Ice Vendors or Entertainers?
### Assessing Media Influence in Contemporary News Environments

<br></br>

### Nicolai Berk<sup>*</sup>

Presentation at the Unanonymous Negroni Drinkers Association

`r Sys.Date()`

.left[.footnote[<sup>*</sup> Dynamics RTG & Humboldt Universität Berlin]]

---

class: middle, center, inverse

# Can a single media outlet still shape opinions in contemporary news environments?

---

# Conflicting theoretical views I

### Emphasis Framing

- Issue attitudes dependent on considerations in mind `r Cite(bib, "Druckman2001")`.
- These considerations can be altered by the the empasis of different frames in the news media `r Cite(bib, "Nelson1997")`.
- As a result, issue attitudes can be manipulated.

___

__Emphasis framing hypothesis__:

.center[*If a newspaper emphasises more positive/negative considerations of an issue, the readers’ attitude about the issue becomes more positive/negative*.]

---

# Conflicting theoretical views II

### Selective Exposure

- Modern news environments offer plenty of choice to consumers `r Cite(bib, c("Arceneaux2013", "Bennett2008"))`.
- Viewers seek out attitude-congruent news content `r Cite(bib, c("Iyengar2009", "Taber2006"))`.
- Hence they will change the news outlet if their preferred news outlet changes its slant.

___

__Opt-out hypothesis__: If a newspaper emphasises more positive/negative considerations of an issue, attitude-incongruent readers will stop consuming the newspaper.

---

# Conflicting theoretical views III

### Selective Exposure II

- However, unclear whether consumers seek out different articles or outlets `r Cite(bib, c("Lau2021", "Marquart2016", "Zillich2021"))`.

___

__Selective avoidance hypothesis__: If a newspaper emphasises more positive/negative considerations of an issue, attitude-incongruent readers will continue consuming the newspaper.

---

class: middle, center, inverse

# The Case

---

# Bild Chief Editor Julian Reichelt

- Germany 2015
- declining readership`
- insert quote on lost readership

<!-- [insert graph of declining readership] -->

<!-- [insert Reichelt picture, maybe Diekmann on the other side of the graph] -->

---

class: middle, center, inverse

# Data

---

# Measurement of News Framing

- BERT for selection of news content about migration
- STM for relative considerations - focus on crime content

---

# Migration Attitude and Readership

- GLES Panel Data preceding the German election 2017

[insert MIP graph]

---

class: middle, center, inverse

# Study Design

---

# Study I

Does a change in migration framing lead to changing migration attitudes?

1. Aggregate-level DiD design exploiting change in Bild migration framing.
2. Assessing all changes in migration framing in the major German dailys.
  1. Aggregate DiD-correlations of frame change and opinion change 
  2. Individual-level estimation of news diet and correlation with over-time opinion changes.

---

# Study 2

Does a change in migration framing lead to changes in readership composition?

Assessment of association of initial and current migration attitude with reported readership.

---

class: middle, center, inverse

# Results

---

# Treatment: Change in Bild Migration Framing

```{r treatment, fig.align='center', dev = 'svg', fig.height = 4, fig.width = 12}

library(here)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(lubridate)

## frames over time
load(here("data/media_daily_2021-11-19.Rdata"))

framing_sum <- 
  merged_media %>% 
  mutate(
    date = floor_date(as.Date(date_new), 'month')
  ) %>% 
  select(-date_new) %>% 
  group_by(paper, date) %>% 
  summarise_all(mean, na.rm = T) %>% 
  pivot_longer(cols = !c(date, paper, crime, n_tot, 
                         n_mig_fra, n_mig_sal, share_mig,
                         medit, deport, refnums, camps, 
                         labmar, capcrime), 
               names_to = 'frame', values_to = 'attention') %>% 
  mutate(paper = recode(paper, 
                        'bild' = 'Bild', 'faz' = 'FAZ', 'spon' = 'Spiegel',
                        'sz' = 'SZ', 'taz' = 'TAZ', 'welt' = 'Welt'
                        )) %>% 
  mutate(frame = recode(frame,
                        'camps_share' = 'Camps', 'crime_share' = 'Petty Crime',
                        'capcrime_share' = 'Capital Crime', 'deport_share' = 'Deportations',
                        'labmar_share' = 'Labour Market', 'medit_share' = 'Mediterranean',
                        'refnums_share' = 'Refugee Numbers'
                        ))
framing_sum$post <- framing_sum$date > as.Date("2017-02-01")

# effect of Diekmann leaving
framing_sum %>%
  filter(date < as.Date('2020-01-01')) %>%
  filter(paper %in% c("Bild", "Welt", "FAZ")) %>% 
  filter(frame %in% c("Camps", "Capital Crime", "Labour Market", "Petty Crime")) %>%
  group_by(paper, frame) %>% 
  ggplot(aes(x = date, y = attention, col = paper)) +
  geom_point(alpha = 0.3) +
  geom_smooth(aes(col = paper, lty = post)) +
  geom_vline(xintercept = as.Date('2017-01-31'), lty = 2, col = 'red') +
  facet_wrap(~frame, scales = "free_y") +
  xlab('')+
  theme_minimal()

```

---

# Study 1: Effects on Migration Attitudes - Case findings

![](vis/case_multiple_dvs.png)

---

# Study 1: Effects on Migration Attitudes - Overall findings

![](vis/frame_effects_did.png){height=50%}

---

class: inverse

### Finding:

- No association of migration framing and change in migration attitude!
- Instead change in media consumption in response to changes in issue framing?

---

# Study 2: Effects on readership

![](vis/reader_share_immatt.png){height=50%}

---

![](vis/imm_bildread_waves.png){height=50%}

---

class: inverse

### Finding:

- No association of migration framing and readership either.

---

class: middle, center, inverse

# Conclusion

---

# Conclusion

- No effects of substantial change in migration framing on issue attitudes.
- No association of changes in migration framing generally with changing issue attitudes.
- No association of sustantial change in migration framing with changing readership behaviour.

__$\rightarrow$ Limited influence of single outlets in diverse media markets!__

---

# Limitations

- Unclear how this translates from tabloids to broadsheets.
- Unclear if salience of migration issue contributed to fixed issue positions.

---

class: middle, center, inverse

# Negroni time!

![](Negroni.png)

---

# Resources

```{r biblio, echo=F, results='asis'}

PrintBibliography(bib)

```