### 2022-06-28 pres/20220217/Thomas_20220128.Rmd

---
title: "Assessing media influence in contemporary news environments"
subtitle: "Down the rabbit hole of conditioning on prior attitude"
author: "Nicolai Berk"
institute: "Humboldt Universität Berlin"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    seal: false
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
    includes:
      after_body: insert-logo.html
---

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#03396d", background_color = "#bfbfbf",
  base_font_size = "25px",
  header_h1_font_size = "2rem",
  header_h2_font_size = "1.75rem",
  header_h3_font_size = "1.25rem",
  header_font_google = google_font("Montserrat", 'Thin 100'),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono"),
  footnote_font_size = "0.75rem"
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = F, warning = F, dev.args = list(bg = 'transparent'), cache = T, out.width="100%", fig.height=6, fig.width=15)

library(tidyverse)
library(dplyr)
library(here)
library(data.table)
library(fixest)
library(lubridate)
library(patchwork)

# load data ####
merged_data <- fread(here('data/merged_2022-02-22.csv'))

# define modeling function

modeltable <- function(data, iv_name, dv_name = "1130_clean") {

  # if(!exists(iv)){stop("Independent not specified.")}

  modeldata <- 
    data %>% 
    mutate(dv = get(dv_name), iv = get(iv_name)) %>% 
    arrange(lfdn, date_clean.x) %>% 
    group_by(lfdn) %>% 
    mutate(dv_lag = lag(dv))

  m_naive <- lm(dv ~ iv, data = modeldata)
  m_dvlag <- lm(dv ~ iv + lfdn, data = modeldata)
  m_wavef <- feglm(dv ~ iv + lfdn | wave, data = modeldata, cluster = c("lfdn", "wave"))
  m_inter <- feglm(dv ~ iv * imm_att | wave, data = modeldata, cluster = c("lfdn", "wave"))

  estimate_table <- 
    data.frame(
      estimate = c("Frame exposure", "CI", "individual FEs", "wave FEs", "Frame X moderate", "CI", "Frame X liberal", "CI"),
      naive    = c(round(m_naive$coefficients[["iv"]], 3), 
                   paste0("[", round(confint(m_naive)["iv",][[1]], 3), ";", 
                          round(confint(m_naive)["iv",][[2]], 3), "]"),
                   "No", "No", "-", "-", "-", "-"),
      lagged   = c(round(m_dvlag$coefficients[["iv"]], 3),
                   paste0("[", round(confint(m_dvlag)["iv",][[1]], 3), ";", 
                          round(confint(m_dvlag)["iv",][[2]], 3), "]"), 
                   "Yes", "No", "-", "-", "-", "-"),
      wavefe   = c(round(m_wavef$coefficients[["iv"]], 3), 
                   paste0("[", round(confint(m_wavef)["iv",][[1]], 3), ";", 
                          round(confint(m_wavef)["iv",][[2]], 3), "]"), 
                   "Yes", "Yes", "-", "-", "-", "-"),
      inter    = c(round(m_inter$coefficients[["iv"]], 3), 
                   paste0("[", round(confint(m_inter)["iv",][[1]], 3), ";", 
                          round(confint(m_inter)["iv",][[2]], 3), "]"),  
                   "No", "Yes", 
                   round(m_inter$coefficients[["iv:imm_attmoderate"]], 3),
                   paste0("[", round(confint(m_inter)["iv:imm_attmoderate",][[1]], 3), ";", 
                          round(confint(m_inter)["iv:imm_attmoderate",][[2]], 3), "]"), 
                   round(m_inter$coefficients[["iv:imm_attliberal"]], 3), 
                   paste0("[", round(confint(m_inter)["iv:imm_attliberal",][[1]], 3), ";", 
                          round(confint(m_inter)["iv:imm_attliberal",][[2]], 3), "]"))
    )

  return(estimate_table)

}

```

```{r bib, include=F, warning = F}
# library(RefManageR)
# BibOptions(check.entries = FALSE, bib.style = "authoryear", style = "text")
# file <- system.file("Bib", "PhD-Proposal.bib", package = "RefManageR")
# bib <- ReadBib("PhD-Proposal.bib", check = F)

```

class: inverse, center, middle

## Does framing matter for news consumption?

---

## Reichelt takeover

More Crime framing of migration in Bild

```{r case_crime}

## define survey dates
gles_p_long <- 
  fread(here('data/gles/Panel/long_cleaned.csv'))

survey_dates <- 
  gles_p_long %>% 
  filter(!is.na(date_clean)) %>% 
  filter(!is.na('1130_clean') & !is.na('1661a_clean')) %>%  
  filter(date_clean < as.Date("2018-01-01")) %>% 
  group_by(wave) %>% 
  summarise(date = max(date_clean)) %>% 
  ungroup() %>% 
  select(date)

## load & transform data frame attention
load(here("data/media_daily_2022-02-02.Rdata"))

merged_media <- 
  merged_media %>% 
  mutate(paper = recode(paper, 
                        'bild' = 'Bild', 'faz' = 'FAZ', 'spon' = 'Spiegel',
                        'sz' = 'SZ', 'taz' = 'TAZ', 'welt' = 'Welt'),
         post = date_new > as.Date("2017-02-01"),
         date_month = floor_date(as.Date(date_new), 'month'))

## visualise
trendplot <- 
  merged_media %>% 
  filter(date_new < as.Date("2018-01-01"), date_new >= as.Date("2016-01-01")) %>% 
  mutate(
    date_month = floor_date(as.Date(date_new), 'month')
  ) %>%
  group_by(paper != "Bild", date_month) %>%
  summarise(across(share_mig:hungary_referendum_share, mean, na.rm = T),
            across(n_mig_sal:n_tot, sum, na.rm = T)) %>%
  mutate(crime_share = crime_schlepp_share + sexual_assault_share,
         post = date_month >= as.Date("2017-02-01")) %>% 
  filter(date_month < as.Date("2020-01-01")) %>% 
  ggplot(aes(x = date_month, y = crime_share, col = `paper != "Bild"`)) +
  geom_line() +
  geom_rug(data = survey_dates, aes(x = date),sides = 't', inherit.aes = F) +
  # geom_point() +
  # geom_smooth(alpha = 0.1) +
  geom_vline(xintercept = as.Date("2017-02-01"), lty = 2, col = "red") +
  ggtitle("Monthly Share of Migration Content Devoted to Crime Frames", "Bild vs. other major daily newspapers, 2016-2017") + 
  xlab("Date") + ylab("Share of Migration Content") +
  # facet_wrap(~paper)
  theme_minimal() +
  theme(legend.position = "none")

# DiD with placebo papers, 2016-17
merged_media <- 
  merged_media %>% 
  filter(date_new < as.Date("2018-01-01"),
         date_new >= "2016-01-01")

treatment_ests <- 
  data.frame()

merged_media$dv    <-  merged_media[["crime_schlepp_share"]] + merged_media[["sexual_assault_share"]] + merged_media[["bild_footer_share"]]
merged_media$post  <-  merged_media$date_new >= as.Date("2017-02-01")

for (p in unique(merged_media$paper)){

  merged_media$treatment <-  merged_media$paper == p

  did_model <- lm(dv ~ post*treatment, data = merged_media)

  est <- did_model$coefficients["postTRUE:treatmentTRUE"][[1]]
  lower <- confint(did_model)["postTRUE:treatmentTRUE", 1]
  upper <- confint(did_model)["postTRUE:treatmentTRUE", 2]

  treatment_ests <- 
    rbind(treatment_ests,
          data.frame(
            model = "DiD",
            paper = p,
            est   = est,
            lower = lower,
            upper = upper
          ))
}

## vis
treatment_ests$paper <- 
  factor(treatment_ests$paper,
          levels = treatment_ests %>%
                    filter(model == "DiD") %>% 
                    arrange(est) %>% 
                    select(paper) %>%
                    unlist()
    )

placebo_papers <- 
  treatment_ests %>% 
  ggplot(aes(x = est, xmin = lower, xmax = upper, 
             y = paper, col = paper != "Bild")) +
  geom_pointrange() +
  geom_vline(xintercept = 0,
             col = "red", lty = 2) +
  theme_minimal() +
  xlab("Estimated DiD") + ylab("") +
  # ggtitle("Papers' Change in Crime Framing (DiD), 2016-2017") +
  theme(legend.position = "None") +
  coord_flip()

trendplot + placebo_papers

```

---

## Case estimate

```{r did_case}

## load estimates
gles_p_long_nona <- fread(file = here('data/gles/Panel/long_cleaned_nona.csv'))

## define post-period
gles_p_long_nona$post <- 
  gles_p_long_nona$date_clean >= as.Date("2017-02-01")

## define treatment: read Bild or not
gles_p_long_nona$treat <- 
  gles_p_long_nona$`1661a_bin` == T

# estimation function ####

did_table <- function(data, dv_name = "1130_clean") {

  ## define dv
  data$dv <- data[[dv_name]]

  data <- 
    data %>% 
    filter(!is.na(dv), !is.na(treat), !is.na(post), !is.na(imm_att))

  m_naive <- feglm(dv ~ treat*post | lfdn, cluster = c("lfdn", "wave"), data = data)
  m_cntrl <- feglm(dv ~ treat*post + imm_att, cluster = c("lfdn", "wave"), data = data)
  m_inter <- feglm(dv ~ treat*post*imm_att, cluster = c("lfdn", "wave"), data = data) # slightly significant difference of liberals
  # m_inter_fe <- feglm(dv ~ treat*post*imm_att, cluster = c("lfdn", "wave"), data = data) # slightly
  #

  estimate_table <- 
      data.frame(
        estimate = c("DiD", "CI", "initial opinion", "individual FEs", "clustered SEs", "DiD X moderate", "CI", "DiD X liberal", "CI"),
        fe    = c(round(m_naive$coefficients[["treatTRUE:postTRUE"]], 3), 
                     paste0("[", round(confint(m_naive)["treatTRUE:postTRUE",][[1]], 3), ";", 
                            round(confint(m_naive)["treatTRUE:postTRUE",][[2]], 3), "]"),
                     "No", "Yes", "Yes", "-", "-", "-", "-"),
        cntrl    = c(round(m_cntrl$coefficients[["treatTRUE:postTRUE"]], 3), 
                     paste0("[", round(confint(m_cntrl)["treatTRUE:postTRUE",][[1]], 3), ";", 
                            round(confint(m_cntrl)["treatTRUE:postTRUE",][[2]], 3), "]"),
                     "Yes", "No", "Yes", "-", "-", "-", "-"),
        inter    = c(round(m_inter$coefficients[["treatTRUE:postTRUE"]], 3), 
                     paste0("[", round(confint(m_inter)["treatTRUE:postTRUE",][[1]], 3), ";", 
                            round(confint(m_inter)["treatTRUE:postTRUE",][[2]], 3), "]"),  
                     "Yes", "No", "Yes", 
                     round(m_inter$coefficients[["treatTRUE:postTRUE:imm_attmoderate"]], 3),
                     paste0("[", round(confint(m_inter)["treatTRUE:postTRUE:imm_attmoderate",][[1]], 3), ";", 
                            round(confint(m_inter)["treatTRUE:postTRUE:imm_attmoderate",][[2]], 3), "]"), 
                     round(m_inter$coefficients[["treatTRUE:postTRUE:imm_attliberal"]], 3), 
                     paste0("[", round(confint(m_inter)["treatTRUE:postTRUE:imm_attliberal",][[1]], 3), ";", 
                            round(confint(m_inter)["treatTRUE:postTRUE:imm_attliberal",][[2]], 3), "]"))
        # inter_fe    = c(round(m_inter$coefficients[["treatTRUE:postTRUE"]], 3), 
        #              paste0("[", round(confint(m_inter)["treatTRUE:postTRUE",][[1]], 3), ";", 
        #                     round(confint(m_inter)["treatTRUE:postTRUE",][[2]], 3), "]"),  
        #              "No", "Yes", 
        #              round(m_inter_fe$coefficients[["treatTRUE:postTRUE:imm_attmoderate"]], 3),
        #              paste0("[", round(confint(m_inter_fe)["treatTRUE:postTRUE:imm_attmoderate",][[1]], 3), ";", 
        #                     round(confint(m_inter_fe)["treatTRUE:postTRUE:imm_attmoderate",][[2]], 3), "]"), 
        #              round(m_inter_fe$coefficients[["treatTRUE:postTRUE:imm_attliberal"]], 3), 
        #              paste0("[", round(confint(m_inter_fe)["treatTRUE:postTRUE:imm_attliberal",][[1]], 3), ";", 
        #                     round(confint(m_inter_fe)["treatTRUE:postTRUE:imm_attliberal",][[2]], 3), "]"))
      )

  return(estimate_table)
}

did_all <- did_table(gles_p_long_nona)

knitr::kable(did_all)

```

### No direct effect, but liberals become more conservative
But could this just be regression to the mean? (And why is the direct effect negative?)

---

## Compared to readers of other conservative outlets

```{r did_case_cons}

# restrict to conservative NP readers
gles_cons <- 
  gles_p_long_nona %>% 
  filter(`1661a_bin` | `1661c_bin` | `1661f_bin`)

did_cons <- did_table(gles_cons)

knitr::kable(did_cons)

rm(gles_p_long_nona, gles_cons)

```

### Liberal effect persists on top of positive effect for conservative bild readers.

---

## What about AfD support?

```{r}

gles_p_long$afd_vote <- gles_p_long$`190b_clean` == "AfD"
## define post-period
gles_p_long$post <- 
  gles_p_long$date_clean >= as.Date("2017-02-01")

## define treatment: read Bild or not
gles_p_long$treat <- 
  gles_p_long$`1661a_bin` == T

knitr::kable(did_table(gles_p_long, dv = "afd_vote"))

```

### Unaffected, if anything, negative.

---

## AfD scalometer

```{r}
knitr::kable(did_table(gles_p_long, dv = "430i_clean"))

```
Seems to *decrease*, at least among liberal and moderate bild readers! But unclear if liberal decrease significantly different from zero.

---

## Takeaways

- Predisposition matters (Zaller, Lazarsfeld)
- But does this generalise beyond this case?

---

class: center, inverse, middle

# Individual estimation of news diet

---

## Individual models

- Estimate individual frame exposure based on news diet
- allows to correlate opinions and migration framing in read newspapers directly

---

## Individual models: Crime

```{r crime_effs}

# IMPORTANT: all non-readers kicked out here! #

# define treatment
merged_data <- 
  merged_data %>% 
  mutate(crime_share = crime_schlepp_share + sexual_assault_share,
         imm_att = factor(imm_att, levels = c("conservative", "moderate", "liberal", "")))

merged_data$n_articles[is.na(merged_data$n_articles)] <- 0

# select 7d lag
merged_data_7d <- 
  merged_data %>% 
  filter(lag == "7") %>% 
  filter(imm_att != "")

rm(merged_data)

crime_effects <- modeltable(merged_data_7d, "crime_share")

knitr::kable(crime_effects)

```

---

## Individual models: Crime II

```{r crime_vis}

merged_data_7d %>% 
  ggplot(aes(crime_share, `1130_clean`)) +
  geom_density_2d_filled() +
  geom_smooth(method = "lm") +
  facet_grid(~imm_att) +
  theme_minimal()

```

## Can this be explained by differences in values?

### The country must be safe

```{r}

merged_data_7d$dv <- merged_data_7d$`1130_clean`
merged_data_7d$iv <- merged_data_7d$crime_schlepp_share
merged_data_7d$security <- as.factor(merged_data_7d$`3320m_clean`)
modeldata <- 
  merged_data_7d %>% 
  filter(!is.na(dv), !is.na(iv), !is.na(security)) %>% 
  select(dv, iv , wave, security, lfdn)

feglm(dv ~ iv * security | wave, data = modeldata, cluster = c("lfdn", "wave"))

```

---

## Individual Model: Crime on AfD support

```{r}
crime_effects_afd <- modeltable(merged_data_7d, "crime_share", dv_name = "430i_clean")

knitr::kable(crime_effects_afd)
```

Positive direct effect, gets weird with interaction (might als not make that much sense to interact)

---

```{r}
merged_data_7d %>% 
  ggplot(aes(crime_share, `430i_clean`)) +
  geom_density_2d_filled() +
  geom_smooth(method = "lm") +
  facet_grid(~imm_att) +
  theme_minimal()
```

## Crime effect on AfD-vote

```{r}
merged_data_7d$afd_vote <- merged_data_7d$`190b_clean` == "AfD"

crime_effects_afd <- modeltable(merged_data_7d, "crime_share", dv_name = "afd_vote")

knitr::kable(crime_effects_afd)
```

No significant effects on AfD support

---

## Does crime effect on attitude persist when controlling for changing issue definitions?

```{r issdef}

merged_data_7d$dv <- merged_data_7d$`1130_clean`
merged_data_7d$crime_def <- 
  (merged_data_7d$`840_c1` == "3411") |
  (merged_data_7d$`840_c2` == "3411") |
  (merged_data_7d$`840_c3` == "3411") |
  (merged_data_7d$`840_c4` == "3411") |
  (merged_data_7d$`840_c5` == "3411")

feglm(dv ~ crime_share*imm_att | wave, cluster = c("lfdn", "wave"), data = merged_data_7d)
feglm(dv ~ crime_share*imm_att + crime_def | wave, cluster = c("lfdn", "wave"), data = merged_data_7d)

```

### Yes.

---

## Individual models: Camps

```{r camp_effs}

# IMPORTANT: all non-readers kicked out here! #

merged_data_7d <- 
  merged_data_7d %>% 
  mutate(camps_share = camps_greece_share + ME_camps_share)

camps_effects <- modeltable(merged_data_7d, "camps_share")

knitr::kable(camps_effects)

```

---

## Individual Models: Camps II

```{r camp_vis}

merged_data_7d %>% 
  ggplot(aes(camps_share, `1130_clean`)) +
  geom_density_2d_filled() +
  geom_smooth(method = "lm") +
  facet_grid(~imm_att)

```

---

## Individual Models: Mediterranean

```{r med_effs}

# IMPORTANT: all non-readers kicked out here! #

merged_data_7d <- 
  merged_data_7d %>% 
  mutate(medit_share = mediterranean_share+ sr_medit_share)

medit_effects <- modeltable(merged_data_7d, "medit_share")

knitr::kable(medit_effects)

```

---

## Individual Models: Refugee numbers

```{r nums_effs}

# IMPORTANT: all non-readers kicked out here! #
nums_effects <- modeltable(merged_data_7d, "refugee_nums_share")

knitr::kable(nums_effects)

```

---

## Salience

Effect of salience not uniform $\rightarrow$ polarising!

```{r salience}

sal_effects <- modeltable(merged_data_7d, "n_articles")

knitr::kable(sal_effects)

```

---

## Crime frame $\rightarrow$ issue definitions

### Tiny positive effect on mentioning migrant crime as most important topic

```{r mip_migrant_crime}

merged_data_7d <- 
  merged_data_7d %>% 
  mutate(dv = `840_c1` == "3411",
         dv_lag = lag(dv),
         ctrl = `840_c1` %in% c(3412:3413, 3750:3759),
         ctrl_lag = lag(ctrl))

model <- feglm(dv ~ crime_schlepp_share + dv_lag | wave, data = merged_data_7d)

knitr::kable(model$coeftable)

```

### ...but stronger effect on other migration topics

```{r mig_oth}

model <- feglm(ctrl ~ crime_schlepp_share + ctrl_lag | wave, data = merged_data_7d %>% filter(imm_att != "")) # stronger positive effect

knitr::kable(model$coeftable)

cor_sal_att <- cor(merged_data_7d$`1140_clean`, merged_data_7d$`1130_clean`, use = "pairwise.complete.obs")

```

---

## Take-Away

- Each frame uniquely shapes public opinion.
- Effect is dependent on existing attitude.
- __Content__ is crucial, salience alone does not explain shifts!
- Public opinion is still a "product of information and predisposition" (Zaller 1992)

## Next steps

- add hypothesis 1b) (interaction with predispositions)
- add analysis of reaction to contrasting frames $\rightarrow$ higher likelihood to stop reading? (sufficient with only crime frame?)
- Frame exposure might affect issue definitions - another paper?

## Addendum: Interaction model with salience

- Correlation of saleince and position on migration: `r cor_sal_att`.
- Quadratic relationship: those with extreme attitudes care about the issue!

```{r sal_att_plot}

merged_data_7d %>% 
  filter(!is.na(`1130_clean`) & !is.na(`1140_clean`)) %>% 
  ggplot(aes(x = `1140_clean`, y = `1130_clean`)) +
  geom_violin(aes(x = as.factor(`1140_clean`))) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2))

```

```{r sal_inter_1}

merged_data_7d <- 
  merged_data_7d %>% 
  mutate(dv = `1130_clean`,
         dv_lag = lag(dv))

# interaction with issue salience
feglm(dv ~ crime_schlepp_share*`1140_clean` | wave, data = merged_data_7d, cluster = c("lfdn", "wave")) # no interaction effect of salience

```

```{r sal_inter_2}

# interaction with issue salience
feglm(dv ~ crime_schlepp_share*`1140_clean` + dv_lag | wave, data = merged_data_7d, cluster = c("lfdn", "wave")) # no interaction effect of salience

```

---

```{r sal_inter_3}

# interaction with issue salience
feglm(dv ~ crime_schlepp_share*as.factor(`1140_clean`) + dv_lag | wave, data = merged_data_7d, cluster = c("lfdn", "wave")) # no interaction effect of salience

cor_lr_att <- cor(merged_data_7d$`1500_clean`, merged_data_7d$`1130_clean`, use = "pairwise.complete.obs")

```

### $\rightarrow$ Interaction of past attitude change not explained by issue attitudes!

---

## Are those with liberal migration attitudes more responsive?

Correlation of general left-right self-placement with migration attitude `r cor_lr_att`.

```{r lr}

feglm(dv ~ crime_schlepp_share*`1500_clean` + dv_lag | wave, data = merged_data_7d, cluster = c("lfdn", "wave")) # no interaction effect of general l-r

```
## Nope.

---

## Are those with liberal migration attitudes more reactive to all frames?

```{r frame_effects, fig.height = 10}

est_ov <- data.frame()

for (treat in merged_data_7d %>% select(n_articles:hungary_referendum_share, -essener_tafel_share) %>% colnames()){
  merged_data_7d$treatment <- merged_data_7d[, get(treat)]
  ests <- feglm(dv ~ treatment*imm_att | wave, data = merged_data_7d, cluster = c("lfdn", "wave"))

  data.frame(
    treatment   = treat,
    treat_est   = ests$coefficients[["treatment"]],
    treat_lower = confint(ests)["treatment",][[1]],
    treat_upper = confint(ests)["treatment",][[2]],
    iamod_est   = ests$coefficients[["treatment:imm_attmoderate"]],
    iamod_lower = confint(ests)["treatment:imm_attmoderate",][[1]],
    iamod_upper = confint(ests)["treatment:imm_attmoderate",][[2]],
    ialib_est   = ests$coefficients[["treatment:imm_attliberal"]],
    ialib_lower = confint(ests)["treatment:imm_attliberal",][[1]],
    ialib_upper = confint(ests)["treatment:imm_attliberal",][[2]]
  ) %>% 
  rbind(est_ov, .) -> est_ov

}

est_ov %>% 
  pivot_longer(cols = treat_est:ialib_upper, names_to = c("estimate", ".value"), names_pattern = "(.*)_(.*)") %>% 
  ggplot(aes(x = est, xmin = lower, xmax = upper, y = estimate, col = estimate == "treat")) +
  geom_pointrange() +
  geom_vline(xintercept = 0, col = "red", lty = 2) +
  facet_wrap(~treatment, scales = "free_x")

```

### No. Frame content matters.

---

## Table of frame effects

```{r}

ov_table <- 
  data.frame(
    c("Asylstreit           ", "None"),
    c("Asylum Law           ", "- All"),
    c("Asylum Reform        ", "- All"),
    c("Australia            ", "- All"),
    c("Balcan               ", "None"),
    c("BAMF                 ", "None"),
    c("Bild Footer          ", "+ Libs"),
    c("Boat full            ", "- All"),
    c("Calais               ", "- Libs"),
    c("Camps Greece         ", "(- Libs)"),
    c("Carneval             ", "None"),
    c("Children             ", "None"),
    c("Church               ", "- All"),
    c("Crime Schlepp        ", "+ Liberals"),
    c("CSU Poverty          ", "POLARISATION"),
    c("Cultural Integration ", "None"),
    c("Deportations         ", "None"),
    c("Economy              ", "(+ Liberals)"),
    c("English              ", "None"),
    c("Episodic             ", "None"),
    c("EU Asylum            ", "- All"),
    c("EU Borders           ", "(- Libs)"),
    c("Families Episodic    ", "(+ Liberals)"),
    c("Finance Bonds        ", "+ All"),
    c("Great Coalition      ", "None"),
    c("Hamburg              ", "- Libs"),
    c("History              ", "- Libs"),
    c("Hungary Referendum   ", "- Libs"),
    c("International Distr. ", "- All"),
    c("Islamic State        ", "Polarisation"),
    c("Interviews Morals    ", "None"),
    c("Jobmarket Integration", "Polarisation"),
    c("Länder               ", "- All"),
    c("Lageso               ", "- Libs"),
    c("Legal Integration    ", "- All"),
    c("Local Distribution   ", "Polarisation/+Cons"),
    c("MENA Camps           ", "None"),
    c("Mexico               ", "None"),
    c("Salience             ", "Polarisation/+Cons"),
    c("NY Stocks            ", "+ All"),
    c("Party Politics       ", "- All"),
    c("Politicians EU       ", "None"),
    c("Polls                ", "None"),
    c("Pop. Statistics      ", "None"),
    c("Refugee cost         ", "None"),
    c("Refugee numbers      ", "None"),
    c("Refugees Berlin      ", "(- Libs)"),
    c("RR protests          ", "None"),
    c("RW Intellectuals     ", "- All"),
    c("Sexual Assault       ", "None"),
    c("Spiegel              ", "- All"),
    c("SR Mediterranean     ", "- All"),
    c("Stocks Asia          ", "+ All"),
    c("Stocks CEE           ", "+ All"),
    c("Stock DAX            ", "+ All"),
    c("Swiss Initiative     ", "None"),
    c("Trump                ", "- All"),
    c("UN Migration Pact    ", "None"),
    c("War Syria            ", "- All/(- Cons)")
  ) %>% t() %>% 
  as.tibble() %>% 
  filter(V2 != "None") %>% 
  arrange(V2)

knitr::kable(ov_table)

```

---