### 2022-06-28 pres/20220225-MethodIssues/HowDoIEstimateThat.Rmd

---
title: "How do I estimate that?"
subtitle: "The perils of panel data"
author: "Nicolai Berk"
institute: "Humboldt Universität Berlin"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    seal: false
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
    includes:
      after_body: insert-logo.html
---

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#03396d", background_color = "#bfbfbf",
  base_font_size = "35px",
  header_h1_font_size = "1.5rem",
  header_h2_font_size = "1.25rem",
  header_h3_font_size = "1.1rem",
  header_font_google = google_font("Montserrat", 'Thin 100'),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono"),
  footnote_font_size = "0.75rem"
)

```

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, message = F, warning = F, cache = T, out.width="75%", fig.height=4, fig.width=10, fig.align='center', eval = T)
knitr::opts_knit$set(root.dir = 'C:/Users/nicol/Dropbox/PhD/Papers/Bild/github/bild')

library(tidyverse)
library(dplyr)
library(data.table)
library(fixest)
library(lubridate)

```

class: inverse, center, middle

# Interactions in panel data and other shenanigans

### Nicolai Berk<sup>*</sup>

Presentation prepared for Will Lowe's Research Design course at Hertie School

`r Sys.Date()`

.left[.footnote[<sup>*</sup> Dynamics RTG & Humboldt Universität Berlin]]

---

class: inverse, middle, center

## The research question

### How do issue definitions in the media affect individual attitudes?

---

## The model

```{r fig.height=5, fig.width=8}
library(dagitty)
dag <- dagitty("dag {
  Information -> Consideration -> Attitude
               }", layout = T)
plot( graphLayout( dag ) )

```

---

## Starting point: Fixed-effect model

$$y_{it} = I_{it} + c_i + c_w + \epsilon_{it}$$

- $y_{it}$ $:=$ attitude of individual $i$ at time $t$
- $I_{it}$ $:=$ Information consumed by individual $i$ at time $t$
- $c_{i}$ $:=$ individual fixed effect (i.e. average individual opinion across time)
- $c_{w}$ $:=$ wave fixed effect (i.e. average opinion across individuals)

---

## However.... (Problem I)

### Zaller: effects of information dependent on

- **individual core values**
  + Depend which considerations are formed in response to information
  + e.g. perceiving a homeless person: bum or unlucky?
- **individual attentiveness to politics**
  + increases receptiveness
  + but also ability to relate information to own values

---

## Modeling

- for direct effect: could just use 2-way FEs (or could I?)
- However, values and attentiveness are (considered to be) **stable across time**
- Hence, including them in the model seems to make FEs rather superfluous

... right?

---

## Candidate I

.left-column[

```{r out.width="100%", fig.height=5, fig.width= 3, fig.align='right'}
library(dagitty)
dag <- dagitty("dag {
  Information -> Attitude
  Information <-  Confounder -> Attitude

               }", layout = T)
plot( graphLayout( dag ) )

```

]

.right-column[

### No individual FEs<sup>1</sup>:

$$y_{it} = I_{it} * V_{i} * A_{i} + c_w + \epsilon_{it}$$

>- Issue: not controlling for time-constant individual differences

.footnote[<sup>1</sup>: Imagine constitutive terms.]

]

---

## Candidate II

### Lagged DV:

$$y_{it} = I_{it} * V_{i} * A_{i} + y_{i, t-1} + c_w + \epsilon_{it}$$

>- Issue: post-treatment bias when it comes to values? (values correlate with dv)

---

## Can PTB be an issue here?

.center[There should be a figure here.]

---

## Candidate III

### With FEs:

$$y_{it} = I_{it} * V_{i} * A_{i} + c_{i} + \epsilon_{it}$$

>- Issue: Collinearity

___

$\rightarrow$ Currently leaning towards first version (no FEs).

---

class: center, middle, inverse

## Suggestions? Hot takes? Emotions?

---

## Problem II: No real control in individual model

### DiD: no problem

```{r case_crimev}

## define survey dates

gles_p_long <- 
  fread('data/gles/Panel/long_cleaned.csv')

survey_dates <- 
  gles_p_long %>% 
  filter(!is.na(date_clean)) %>% 
  filter(!is.na('1130_clean') & !is.na('1661a_clean')) %>%  
  filter(date_clean < as.Date("2018-01-01")) %>% 
  group_by(wave) %>% 
  summarise(date = max(date_clean)) %>% 
  ungroup() %>% 
  select(date)

rm(gles_p_long)

## load & transform data frame attention
load("data/media_daily_2022-02-02.Rdata")

merged_media <- 
  merged_media %>% 
  mutate(paper = recode(paper, 
                        'bild' = 'Bild', 'faz' = 'FAZ', 'spon' = 'Spiegel',
                        'sz' = 'SZ', 'taz' = 'TAZ', 'welt' = 'Welt'),
         post = date_new > as.Date("2017-02-01"),
         date_month = floor_date(as.Date(date_new), 'month'))

## visualise
trendplot <- 
  merged_media %>% 
  filter(date_new < as.Date("2018-01-01"), date_new >= as.Date("2016-01-01")) %>% 
  mutate(
    date_month = floor_date(as.Date(date_new), 'month')
  ) %>%
  group_by(paper != "Bild", date_month) %>%
  summarise(across(share_mig:hungary_referendum_share, mean, na.rm = T),
            across(n_mig_sal:n_tot, sum, na.rm = T)) %>%
  mutate(crime_share = crime_schlepp_share + sexual_assault_share,
         post = date_month >= as.Date("2017-02-01")) %>% 
  filter(date_month < as.Date("2020-01-01")) %>% 
  ggplot(aes(x = date_month, y = crime_share, col = `paper != "Bild"`)) +
  geom_line() +
  geom_rug(data = survey_dates, aes(x = date),sides = 't', inherit.aes = F) +
  # geom_point() +
  # geom_smooth(alpha = 0.1) +
  geom_vline(xintercept = as.Date("2017-02-01"), lty = 2, col = "red") +
  ggtitle("Monthly Share of Migration Content Devoted to Crime Frames", "Bild vs. other major daily newspapers, 2016-2017") + 
  xlab("Date") + ylab("Share of Migration Content") +
  # facet_wrap(~paper)
  theme_minimal() +
  theme(legend.position = "none")

# DiD with placebo papers, 2016-17
merged_media <- 
  merged_media %>% 
  filter(date_new < as.Date("2018-01-01"),
         date_new >= "2016-01-01")

treatment_ests <- 
  data.frame()

merged_media$dv    <-  merged_media[["crime_schlepp_share"]] + merged_media[["sexual_assault_share"]] + merged_media[["bild_footer_share"]]
merged_media$post  <-  merged_media$date_new >= as.Date("2017-02-01")

for (p in unique(merged_media$paper)){

  merged_media$treatment <-  merged_media$paper == p

  did_model <- lm(dv ~ post*treatment, data = merged_media)

  est <- did_model$coefficients["postTRUE:treatmentTRUE"][[1]]
  lower <- confint(did_model)["postTRUE:treatmentTRUE", 1]
  upper <- confint(did_model)["postTRUE:treatmentTRUE", 2]

  treatment_ests <- 
    rbind(treatment_ests,
          data.frame(
            model = "DiD",
            paper = p,
            est   = est,
            lower = lower,
            upper = upper
          ))
}

## vis
treatment_ests$paper <- 
  factor(treatment_ests$paper,
          levels = treatment_ests %>%
                    filter(model == "DiD") %>% 
                    arrange(est) %>% 
                    select(paper) %>%
                    unlist()
    )

placebo_papers <- 
  treatment_ests %>% 
  ggplot(aes(x = est, xmin = lower, xmax = upper, 
             y = paper, col = paper != "Bild")) +
  geom_pointrange() +
  geom_vline(xintercept = 0,
             col = "red", lty = 2) +
  theme_minimal() +
  xlab("Estimated DiD") + ylab("") +
  # ggtitle("Papers' Change in Crime Framing (DiD), 2016-2017") +
  theme(legend.position = "None") +
  coord_flip()

gridExtra::grid.arrange(trendplot, placebo_papers, nrow = 1)

rm(merged_media)

```

---

### Individual estimates: problem?

```{r}

# load data ####
merged_data <- 
  fread('data/merged_2022-02-22.csv') %>% 
  mutate(crime_share = crime_schlepp_share + sexual_assault_share)

merged_data %>% 
  filter(!is.na(crime_share)) %>% 
  ggplot(aes(x = crime_share)) +
  geom_histogram(fill = "lightblue", alpha = 0.7) +
  ggtitle("Individual exposure to crime frame") +
  xlab("Share of crime news out of all news on migration") + ylab("Count")+
  theme_minimal() 
  # facet_wrap(~wave)

```

---

## Potential solutions?

- Create bins for different levels of exposure?
- Actually not a problem to estimate 2-way FE model? (but *reverse causality*?)
- something else?

---

class: center, middle, inverse

## Suggestions? Hot takes? Emotions?

---

## Mo' problems

- DiD standard errors: all solved with bootstrap?
- do 2-way FEs account for switching in DiD groups?
- standard error clustering
- test for endogeneity

---

class: center, middle, inverse

## Like and subscribe for more research riddled with problems