### 2022-06-28 pres/20220506/Dublin.Rmd

---
title: "How the News Media Shapes Public Opinion"
author: "Nicolai Berk"
institute: "Humboldt Universität Berlin"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    seal: false
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
    includes:
      after_body: insert-logo.html
---

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#004A95", background_color = "#bfbfbf",
  base_font_size = "36px",
  header_h1_font_size = "1.3rem",
  header_h2_font_size = "1.2rem",
  header_h3_font_size = "1.05rem",
  header_font_google = google_font("Lato", '700'),
  text_font_google   = google_font("Lato", "400"),
  code_font_google   = google_font("Fira Mono"),
  footnote_font_size = "0.75rem"
)

```

```{r, load_refs, include=FALSE, cache=FALSE}
library(RefManageR)
BibOptions(check.entries = FALSE, 
           bib.style = "authoryear",
           cite.style = "authoryear",
           style = "markdown",
           hyperlink = FALSE,
           dashed = FALSE)
myBib <- ReadBib("C:/Users/nicol/Dropbox/Studium/BibTex/PhD.bib", check = FALSE)
```

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = FALSE, 
  message = F, 
  warning = F, 
  cache = T, 
  out.width="80%", 
  fig.height=9, 
  fig.width=16, 
  fig.align='center', 
  eval = T, 
  dev = "svg"
  )
# knitr::opts_knit$set(root.dir = 'C:/Users/nicol/Dropbox/PhD/Papers/Bild/github/bild')

library(tidyverse)
library(dplyr)
library(data.table)
library(fixest)
library(lubridate)
library(patchwork)

```

class: inverse, center, middle

# Framing Effects in the Wild

## How the News Media Shapes Public Opinion by Defining Political Issues

#### Nicolai Berk<sup>*</sup>

Presentation prepared for the 4th Annual CompText Conference, Dublin

`r Sys.Date()`

.left[.footnote[<sup>*</sup> Dynamics RTG & Humboldt Universität Berlin]]

---

## Migration coverage can look like this

```{r out.width="80%"}

knitr::include_graphics("collage_ukraine.png")

```

---

## ...or like this

```{r, out.width='100%'}

knitr::include_graphics("collage.png")

```

---

## Does News Content Affect Political Attitudes?

<br>

A bullet-point review:

- Experimental research shows large framing effects `r Citep(myBib[c("kahneman2011thinking", "Leeper2020")])`.
- Unclear if externally valid `r Citep(myBib[c("Barabas2010")])`.
- Increasing reliance on **observational studies** and **field experiments**.

---

## Research Gap

<br>

Current focus on:

- **Structural, long-term changes** in media markets `r Citep(myBib[c("Foos2022", "Grossman2022")])`.
- Effects of **changing news outlets** `r Citep(myBib[c("Broockman2022", "Guess2021")])`.
- **Average** treatment effects.

---

## Research Gap II

<br>

Unclear:

- Which news **content** is effective?
- **Everyday, medium-term** impact?
- Effect dependent on individual **predispositions**?

---

## Argument

<br>

1. News media influences public opinion by __emphasising certain considerations__ over others.
2. **Predispositions** structure individual susceptibility.

---

class: center, middle, inverse

# Theoretical Framework

---

## Expectancy-Value Model `r Citep(myBib["Ajzen2000"])`

```{r out.width="70%"}

knitr::include_graphics("CognitiveStorage.png")

```

---

## Theoretical Model

.left-column[

![](minDAG.png)

]

.right-column[

Cognitive **association** with a consideration is a **function of emphasis** of this aspect in media content `r Citep(myBib[c("Leeper2020",  "Nelson1997")])`.

<br>

### H1: Increased **exposure to negative considerations** results in more **negative attitudes** about an issue.

]

---

## Availability

<br>

- Experimental research suggests that **holding a consideration in mind decreases susceptibility to framing** `r Citep(myBib[c("Druckman2012")])`.
- Hence media content only affects opinions when exposed individual **learns something new**.

### H2: Individuals are less susceptible when exposed to considerations that are **already accessible** to them.

---

class: center, middle, inverse

# The Case

---

## Bild's migration coverage following 2015

.left-column[

![Source: Wikipedia](https://upload.wikimedia.org/wikipedia/commons/thumb/3/3c/Kai_Diekmann_MediaMakers.jpg/435px-Kai_Diekmann_MediaMakers.jpg)

![Source: Wikipedia](https://upload.wikimedia.org/wikipedia/commons/thumb/5/5a/2018-11-07-Julian_Reichelt-Maischberger-1141.jpg/800px-2018-11-07-Julian_Reichelt-Maischberger-1141.jpg)

]

.right-column[

- Migration is __the__ big topic preceding the 2017 _Bundestag_ election in Germany.
- Bild editor-in-chief Kai Diekmann left in February 2017, with Reichelt taking over.
- Reichelt increasingly **framed migrants as criminals** `r Citep(myBib[c("Niggemeier2018Bild")])`.

]

---

## Bild's migration coverage following 2015 II

.left-column[

![Source: Wikipedia](https://upload.wikimedia.org/wikipedia/commons/thumb/3/3c/Kai_Diekmann_MediaMakers.jpg/435px-Kai_Diekmann_MediaMakers.jpg)

![Source: Wikipedia](https://upload.wikimedia.org/wikipedia/commons/thumb/5/5a/2018-11-07-Julian_Reichelt-Maischberger-1141.jpg/800px-2018-11-07-Julian_Reichelt-Maischberger-1141.jpg)

]

.right-column[

Interesting because:

- **Clearly identifiable short-term-change** in migration coverage.
- Clear expectations regarding impact.
- Changing editor was not debated in the media $\rightarrow$ __No co-treatment__! `r Citep(myBib["Spirig2020"])`.

]

---
## Bild's migration coverage following 2015 III

.left-column[

![](bild_umgebracht_asyl.jpeg)
![](bild_auslaender.jpeg)

]

.right-column[

**Least likely case**: 
- Issue sorting, 
- pre-treatment, 
- salient party cues, 
- emotional issue.

]

---

class: center, middle, inverse

# Study Design

---

## Measurement: Emphasis Framing in Print Media

<br>

#### Independent Variable

- Collected 2.5M news articles from major German daily newspapers.
- Two-step identification of issue definitions:
  1. Supervised deep-learning model (BERT) to **select migration content**.
  2. Structural Topic Model to **identify considerations**.

---

# Identifying migration content

#### Problem: migration content is rare

### Solution: oversample migration content

1. Define **seed dictionary** of migration terms.
2. Extend dictionary using word embeddings.
3. Oversample from articles mentioning immigration terms.

---

# Identifying migration content II

<br>

- RA annotates 1800 articles.
- Fine-tune **German BERT** to identify migration content (accuracy 0.96, F1: 0.94, recall: 0.93, precision: 0.95).
- Predict label for 2.5M news articles: 90,000 about migration.

---

# Identifying Considerations about Immigration `r Citep(myBib["Nicholson2020"], before = "following ")`

<br>

- Apply **STM** with 60 topics `r Citep(myBib["Roberts2014"])` to identify migration frames.
- **Annotate** all topics, **select** the ones about crime.
- Estimate daily prevalence for each newspaper.

---

## Change in Crime News post-Reichelt

```{r case_crime}

## define survey dates
gles_p_long <- 
  fread('C:/Users/nicol/Dropbox/PhD/Papers/Bild/github/bild/data/gles/Panel/long_cleaned.csv')

survey_dates <- 
  gles_p_long %>% 
  filter(!is.na(date_clean)) %>% 
  filter(!is.na('1130_clean') & !is.na('1661a_clean')) %>%  
  filter(date_clean < as.Date("2018-01-01")) %>% 
  group_by(wave) %>% 
  summarise(date = max(date_clean)) %>% 
  ungroup() %>% 
  select(date)

rm(gles_p_long)

## load & transform data frame attention
load("C:/Users/nicol/Dropbox/PhD/Papers/Bild/github/bild/data/media_daily_2022-02-02.Rdata")

merged_media <- 
  merged_media %>% 
  mutate(paper = recode(paper, 
                        'bild' = 'Bild', 'faz' = 'FAZ', 'spon' = 'Spiegel',
                        'sz' = 'SZ', 'taz' = 'TAZ', 'welt' = 'Welt'),
         post = date_new > as.Date("2017-02-01"),
         date_month = floor_date(as.Date(date_new), 'month'))

## visualise
trendplot <- 
  merged_media %>% 
  filter(date_new < as.Date("2018-01-01"), date_new >= as.Date("2016-01-01")) %>% 
  mutate(
    date_month = floor_date(as.Date(date_new), 'month')
  ) %>%
  group_by(paper != "Bild", date_month) %>%
  summarise(across(share_mig:hungary_referendum_share, mean, na.rm = T),
            across(n_mig_sal:n_tot, sum, na.rm = T)) %>%
  mutate(crime_share = crime_schlepp_share + sexual_assault_share,
         post = date_month >= as.Date("2017-02-01")) %>% 
  filter(date_month < as.Date("2020-01-01")) %>% 
  ggplot(aes(x = date_month, y = crime_share, col = `paper != "Bild"`)) +
  geom_line(size = 2) +
  geom_rug(data = survey_dates, aes(x = date),sides = 't', inherit.aes = F, size = 2) +
  # geom_point() +
  # geom_smooth(alpha = 0.1) +
  geom_vline(xintercept = as.Date("2017-02-01"), lty = 2, col = "red", size = 2) +
  # ggtitle("Monthly Share of Migration Content Devoted to Crime Frames", "Bild vs. other major daily newspapers, 2016-2017") + 
  xlab("") + ylab("Share of Migration Content") +
  # facet_wrap(~paper)
  # theme(legend.position = "none") +
  theme_xaringan() + 
  scale_color_discrete(name = "Outlet", labels = c("Bild", "Other"))

# DiD with placebo papers, 2016-17
merged_media <- 
  merged_media %>% 
  filter(date_new < as.Date("2018-01-01"),
         date_new >= "2016-01-01")

treatment_ests <- 
  data.frame()

merged_media$dv    <-  merged_media[["crime_schlepp_share"]] + merged_media[["sexual_assault_share"]] + merged_media[["bild_footer_share"]]
merged_media$post  <-  merged_media$date_new >= as.Date("2017-02-01")

for (p in unique(merged_media$paper)){

  merged_media$treatment <-  merged_media$paper == p

  did_model <- lm(dv ~ post*treatment, data = merged_media)

  est <- did_model$coefficients["postTRUE:treatmentTRUE"][[1]]
  lower <- confint(did_model)["postTRUE:treatmentTRUE", 1]
  upper <- confint(did_model)["postTRUE:treatmentTRUE", 2]

  treatment_ests <- 
    rbind(treatment_ests,
          data.frame(
            model = "DiD",
            paper = p,
            est   = est,
            lower = lower,
            upper = upper
          ))
}

## vis
treatment_ests$paper <- 
  factor(treatment_ests$paper,
          levels = treatment_ests %>%
                    filter(model == "DiD") %>% 
                    arrange(est) %>% 
                    select(paper) %>%
                    unlist()
    )

placebo_papers <- 
  treatment_ests %>% 
  ggplot(aes(x = est, xmin = lower, xmax = upper, 
             y = paper, col = paper != "Bild")) +
  geom_pointrange(size = 2) +
  geom_vline(xintercept = 0,
             col = "red", lty = 2, size = 2) +
  xlab("Estimated DiD") + ylab("") +
  # ggtitle("Papers' Change in Crime Framing (DiD), 2016-2017") +
  theme(legend.position = "None") +
  coord_flip() + theme_xaringan()

trendplot

# rm(merged_media)

```

---

## Placebo papers

```{r}

placebo_papers

```

---

## Measurement: Immigration Attitude

<br>

#### Dependent Variable

- German Longitudinal Election Study (GLES) Panel
- 7-point Likert-scale asking whether immigration to Germany should be 
  + made easier (-3) or
  + restricted (3).

---

## Measurement: News Consumption and Availability

#### Exposure to Treatment

- Question on newspapers consumed last week.
- Simple binary variable indicating *Bild* consumption.
- Alternative: Calculate individual exposure to crime framing.

--

#### Availability

- **Initial migration opinion**.
- Reporting **migrant crime as MIP**.

---

# Estimation

**Difference-in-Differences (DiD)**

- Comparing Bild readers and non-readers,
- before and after Reichelt takeover.
- Interaction with initial report of migrant crime as most important problem.

$$y_{it} = \alpha + \beta * Post_t *  Treatment_i * y_{i,t=1} + \epsilon_{gt}$$

.footnote[(Constitutive terms omitted for brevity.)]

---

# Estimation II

  **Fixed-effects-model**
- Generate estimates of **individual exposure** to consideration.
- Control for **individual** and **wave fixed effects**.
- Interaction with prior report of migrant crime.

$$y_{it} = \beta * CrimeExposure_{it} * y_{i,t=1} + \rho_i + \phi_t + \epsilon_{it}$$
.footnote[(Constitutive terms omitted for brevity.)]

---

class: center, middle, inverse

# Results

---

# DiD Model

```{r did-data}

source("C:/Users/nicol/Dropbox/PhD/Papers/Bild/github/bild/code/functions.R")

gles_p_long_1617 <- 
  DiD_data_prep()

```

```{r did-ests}

gles_p_long <- fread('C:/Users/nicol/Dropbox/PhD/Papers/Bild/github/bild/data/gles/Panel/long_cleaned.csv')

set.seed(42)
B <- 1000

## no resampling
gles_p_long <- 
  gles_p_long %>% 
  select(lfdn, date_clean, `1130_clean`, wave) %>% 
  filter(!is.na(`1130_clean`), wave == "1") %>% 
  mutate(init_mig = `1130_clean`) %>% 
  select(lfdn, init_mig) %>% 
  right_join(gles_p_long, by = "lfdn")

## with resampling
## still need to think about implications of this
gles_p_long <-
  gles_p_long %>%
  select(lfdn, date_clean, `1130_clean`) %>%
  filter(!is.na(`1130_clean`)) %>%
  arrange(lfdn, date_clean) %>%
  group_by(lfdn) %>%
  slice(1) %>%
  mutate(init_mig_res = `1130_clean`) %>%
  select(lfdn, init_mig_res) %>%
  right_join(gles_p_long, by = "lfdn")

gles_p_long <- 
  gles_p_long %>% 
  mutate(mip_mig_crime_lag = lag(mip_mig_crime)) %>% 
  filter(!is.na(mip_mig_crime)) %>% 
  arrange(lfdn, date_clean) %>% 
  group_by(lfdn) %>% 
  filter(wave == "1") %>% 
  mutate(mip_mig_crime_init = mip_mig_crime) %>% 
  select(lfdn, mip_mig_crime_init) %>% 
  right_join(gles_p_long, by = "lfdn")

gles_p_long <- 
  gles_p_long %>% 
  filter(wave == "1") %>% 
  mutate(bild_init = ifelse(is.na(`1661a_clean`), NA, `1661a_clean` > 0)) %>% 
  select(bild_init, lfdn) %>% 
  right_join(gles_p_long, by = "lfdn")

gles_p_long <- 
  gles_p_long %>% 
  arrange(lfdn, date_clean) %>% 
  group_by(lfdn) %>% 
  slice(1) %>% 
  mutate(treat_alt = ifelse(is.na(`1661a_clean`), NA, `1661a_clean` > 0)) %>% 
  select(treat_alt, lfdn) %>% 
  right_join(gles_p_long, by = "lfdn")

## vis dv by treatment group and conditioning var

did_desc_plot <- 
  gles_p_long %>%
  filter(date_clean < as.Date("2018-01-01")) %>% 
  filter(!is.na(`1130_clean`) & !is.na(bild_init)) %>% 
  mutate(init_mig_cat = 
           case_when(
             init_mig > 0 ~ "01 - Conservative",
             init_mig == 0 ~ "02 - Neutral",
             init_mig < 0 ~ "03 - Liberal"
           )) %>% 
  group_by(bild_init, wave, init_mig) %>% 
  select(`1130_clean`, date_clean) %>% 
  summarise(dv = mean(`1130_clean`, na.rm = T),
            dv_sd = sd(`1130_clean`, na.rm = T),
            date_min = min(date_clean, na.rm = T),
            n_resps = n()) %>%
  mutate(
    lower = dv + qt(0.025, n_resps - 1) * (dv_sd/sqrt(n_resps)),
    upper = dv + qt(0.975, n_resps - 1) * (dv_sd/sqrt(n_resps))
  ) %>% 
  ggplot(aes(date_min, dv, ymin = lower, ymax = upper, col = !bild_init, fill = !bild_init)) +
  geom_line() +
  geom_ribbon(alpha = 0.5) +
  geom_vline(xintercept = as.Date("2017-02-01"), lty = 2, col = "red") +
  facet_grid(~init_mig) +
  ggtitle("Average migration opinion dependent on initial migration opinion", 
          "Bild vs. other major daily newspapers, 2016-2017") + 
  xlab("Date") + ylab("Migration Opinion") +
  theme_minimal() +
  theme(legend.position = "none")

ggsave("C:/Users/nicol/Dropbox/PhD/Papers/Bild/github/bild/paper/vis/did_descriptive.svg",
       did_desc_plot,
       device = "svg", width = 10, height = 6)

gles_p_long_1617 <- 
  gles_p_long %>% 
        filter(date_clean >= as.Date("2016-01-01"),
               date_clean <  as.Date("2018-01-01"))

rm(gles_p_long)

gles_p_long_1617$dv <- gles_p_long_1617$`1130_clean`
gles_p_long_1617$treat <- gles_p_long_1617$bild_init
gles_p_long_1617$post <- gles_p_long_1617$date_clean>= as.Date("2017-02-01")

# model 1: direct effect
ate <- 
  feglm(dv ~ post * treat, 
      cluster = c("lfdn"), 
      data = gles_p_long_1617)

inter <- 
  feglm(dv ~ post * treat * init_mig, 
      cluster = c("lfdn"), 
      data = gles_p_long_1617)

points <- ate$coefficients
ests <- cbind(points, confint(ate))

plot_did_ate <- 
  ests %>%
  filter(row.names(.) == "postTRUE:treatTRUE") %>% 
  ggplot(aes(y = row.names(.), x = points, xmin = `2.5 %`, xmax = `97.5 %`)) +
  geom_pointrange(size = 2) +
  geom_vline(xintercept = 0, col = "red", lty = 2, size = 2) +
  ylab("") + xlab("Estimate") +
  xlim(-0.1, 0.3) + 
  scale_y_discrete(labels = rev(c("DiD"))) +
  ggtitle("Average treatment effect") + 
  theme_xaringan()

points <- inter$coefficients
ests <- cbind(points, confint(inter))

plot_did_inter <- 
  ests %>% 
  filter(row.names(.) %in% c("postTRUE:treatTRUE", "postTRUE:treatTRUE:init_mig")) %>% 
  mutate(name = fct_reorder(row.names(.), desc(row.names(.)))) %>% 
  ggplot(aes(y = name, x = points, xmin = `2.5 %`, xmax = `97.5 %`)) +
  geom_pointrange(size = 2) +
  geom_vline(xintercept = 0, col = "red", lty = 2, size = 2) +
  ylab("") + xlab("Estimate") +
  ggtitle("DiD-estimate with interaction") +
  xlim(-0.1, 0.3) + 
  scale_y_discrete(labels = rev(c("DiD", "DiD X Initial Opinion"))) +
  ggtitle("Heterogenous treatment effects") + 
  theme_xaringan()

did_groups <- 
  gles_p_long_1617 %>% 
  filter(!is.na(treat), !is.na(init_mig)) %>%
  group_by(treat, post, init_mig) %>% 
  summarise(
    dv_point = mean(dv, na.rm = T)
    ) %>% 
  filter(post) %>% 
  pivot_wider(
    names_from = treat,
    values_from = dv_point
  ) %>% 
  mutate(difference = `TRUE` - `FALSE`) %>% 
  ggplot(aes(x = init_mig, y = difference)) +
  geom_col() + 
  theme_xaringan() + xlab("") + ylab("") +
  ggtitle("Effect across initial migration attitude")

load('C:/Users/nicol/Dropbox/PhD/Papers/Bild/github/bild/data/did_bootstraps.RData')

direct <- coef(inter)["postTRUE:treatTRUE"]
interact <- coef(inter)["postTRUE:treatTRUE:init_mig"]

did_bootstrap <- 
  ggplot() +
  geom_histogram(data = gles_p_long_1617 %>% filter(wave == "1"), mapping = aes(x = init_mig, after_stat(density)), alpha = 0.4, bins = 8) +
  geom_abline(data = bs_did_ests, aes(slope = IA, intercept = DiD, group = boots), alpha= 0.05) +
  geom_abline(slope = interact, intercept = direct, col = "red") +
  geom_hline(yintercept = 0, col = "black") +
  scale_x_continuous(limits=c(-3,3)) +
  scale_y_continuous(limits=c(0,.6)) +
  ggtitle(paste0(B, " bootstrap estimates of heterogenous treatment effect")) +
  ylab("Estimate") + xlab("Initial migration attitude") +
  theme_xaringan()

ggsave(plot = did_bootstrap,
       filename = 'C:/Users/nicol/Dropbox/PhD/Papers/Bild/github/bild/paper/vis/DiDXinit_imm.svg', device = "svg")

plot_did_ate / plot_did_inter

```

---

## Conditioning on Availability

```{r did_bs}

did_bootstrap

```

---

## Or, descriptively:

```{r}

did_groups

```

---

# Individual-level exposure model

```{r ind_exp}

merged_data <- fread(file = "C:/Users/nicol/Dropbox/PhD/Papers/Bild/github/bild/data/merged_2022-03-05.csv")

merged_data <- 
  merged_data %>% 
  mutate(crime_share = crime_schlepp_share + sexual_assault_share,
         dv = `1130_clean`) %>% 
  filter(lag == 7)

merged_data <- 
  merged_data %>% 
  filter(!is.na(`1130_clean`)) %>% 
  group_by(lfdn) %>% 
  arrange(lfdn, date_clean.x) %>% 
  select(`1130_clean`, lfdn) %>% 
  slice(1) %>% 
  mutate(init_mig = `1130_clean`) %>%
  select(init_mig, lfdn) %>% 
  right_join(merged_data, by = "lfdn")

# add lagged crime_share and dv
merged_data <- 
  merged_data %>%
  ungroup() %>% 
  arrange(lfdn, date_clean.x) %>% 
  group_by(lfdn) %>% 
  filter(!is.na(crime_share)) %>% 
  mutate(crime_share_lag = lag(crime_share),
         dv_lag = lag(dv))

# estimation ####

# individual exposure ####

# direct crime
ind_ate <- 
  feglm(dv ~ crime_share | wave + lfdn,  # positive, p < 0.1
      cluster = c("lfdn"), data = merged_data)

# crime X initial imm_att
ind_inter <- 
  feglm(dv ~ crime_share:init_mig + crime_share | wave + lfdn,  
      cluster = c("lfdn"), data = merged_data)

ind_inter_ctrl <- 
  feglm(dv ~ crime_share:init_mig + crime_share + mip_mig_crime | wave + lfdn,  
      cluster = c("lfdn"), data = merged_data)

points <- ind_ate$coefficients
ests <- cbind(points, confint(ind_ate))
plot_ind_ate <- 
  ests %>% 
  # mutate(name = factor(row.names(.), levels = row.names(.))) %>% 
  ggplot(aes(y = row.names(.), x = points, xmin = `2.5 %`, xmax = `97.5 %`)) +
  geom_pointrange(size = 2, col = "#004A95") +
  geom_vline(xintercept = 0, col = "red", lty = 2, size = 2) +
  ylab("") + xlab("") +
  scale_y_discrete(limits=rev(row.names(ests)),
                   labels = rev(c("Crime Exposure"))) +
  xlim(-0.75, 2) + 
  theme_xaringan()

points <- ind_inter$coefficients
ests <- cbind(points, confint(ind_inter))
plot_ind_inter <- 
  ests %>% 
  # mutate(name = factor(row.names(.), levels = row.names(.))) %>% 
  ggplot(aes(y = row.names(.), x = points, xmin = `2.5 %`, xmax = `97.5 %`)) +
  geom_pointrange(size = 2, col = "#004A95") +
  geom_vline(xintercept = 0, col = "red", lty = 2, size = 2) +
  ylab("") + xlab("") +
  scale_y_discrete(limits=rev(row.names(ests)),
                   labels = rev(c("Crime Exposure", "Crime Exposure\n X Initial Opinion"))) +
  xlim(-0.75, 2) +
  theme_xaringan()

plot_ind_ate/plot_ind_inter

```

---

# Moderator: MIP

```{r}

# DiD
# model 1: direct effect
ate <- 
  feglm(dv ~ post * treat, 
      cluster = c("lfdn"), 
      data = gles_p_long_1617)

inter <- 
  feglm(dv ~ post * treat * mip_mig_crime_init, 
      cluster = c("lfdn"), 
      data = gles_p_long_1617)

points <- ate$coefficients
ests <- cbind(points, confint(ate))

plot_did_ate <- 
  ests %>%
  filter(row.names(.) == "postTRUE:treatTRUE") %>% 
  ggplot(aes(y = row.names(.), x = points, xmin = `2.5 %`, xmax = `97.5 %`)) +
  geom_pointrange(size = 2) +
  geom_vline(xintercept = 0, col = "red", lty = 2, size = 2) +
  ylab("") + xlab("Estimate") +
  scale_y_discrete(labels = rev(c("DiD"))) +
  ggtitle("Average treatment effect") + 
  theme_xaringan()

points <- inter$coefficients
ests <- cbind(points, confint(inter))

plot_did_inter <- 
  ests %>% 
  filter(row.names(.) %in% c("postTRUE:treatTRUE", "postTRUE:treatTRUE:mip_mig_crime_initTRUE")) %>% 
  mutate(name = fct_reorder(row.names(.), desc(row.names(.)))) %>% 
  ggplot(aes(y = name, x = points, xmin = `2.5 %`, xmax = `97.5 %`)) +
  geom_pointrange(size = 2) +
  geom_vline(xintercept = 0, col = "red", lty = 2, size = 2) +
  ylab("") + xlab("Estimate") +
  ggtitle("DiD-estimate with interaction") +
  scale_y_discrete(labels = rev(c("DiD", "DiD X Initial Opinion"))) +
  ggtitle("Heterogenous treatment effects") + 
  theme_xaringan()

plot_did_ate/plot_did_inter

```

---

# Summary

<br>

- Media **content affects attitude** formation.
- **ATE is null** when not moderating with prior attitude.
- Effect conditioned by **Prior availability** of consideration.

### $\rightarrow$ Attitude is a **product of information and predisposition** `r Citep(myBib["Zaller1992"])`.

---

# Limitations

<br>

#### Unclear:

- How important are these effects in **more fragmented media landscapes**?
- Which content affects **which groups**?
- Which **other actors** can affect framing in the media?

---

class: center, middle, inverse

# Thank you!

---

class: center, middle, inverse

## This research is generously funded by:

```{r out.width="50%"}

knitr::include_graphics("banner_new.png")

```

```{r out.width="50%"}

knitr::include_graphics("dfg_logo_englisch_blau_en.jpg")

```

---

# References

.footnote[

```{r refs, echo=FALSE, results="asis", cache=FALSE}

PrintBibliography(myBib, end = 4)

```

]
---

# References II

.footnote[

```{r refs2, echo=FALSE, results="asis", cache=FALSE}

PrintBibliography(myBib, start = 5, end = 8)

```

]
---

# References II

.footnote[

```{r refs3, echo=FALSE, results="asis", cache=FALSE}

PrintBibliography(myBib, start = 9)

```

]
---

class: center, middle, inverse

# Appendix

---

class: inverse, center, middle

## Variables

---

## Individual Crime Exposure

```{r}

merged_data %>% 
  ggplot(aes(crime_share)) +
  geom_histogram() +
  xlab("Share of migration content consumed concerning crime") +
  theme_xaringan()

```

---

# Dependent Variable

```{r}

gles_p_long_1617 %>% 
  filter(!is.na(`1130_clean`), !
           readership %in% c("", "Other", "None", "FR")) %>% 
  group_by(wave, readership) %>% 
  summarise(dv = mean(`1130_clean`, na.rm = T),
            date_min = min(date_clean)) %>% 
  ggplot(aes(x = date_min, y = dv, col = readership)) +
  geom_line(size = 2) +
  theme_xaringan()

```

---

class: inverse, center, middle

# Mediated by pre-existing issue definitions?

---

## Control for mention of migrant crime as MIP in DiD:

```{r inter_ctrl}

inter_ctrl <- 
  feglm(dv ~ post * treat * init_mig + mip_mig_crime, 
      cluster = c("lfdn"), 
      data = gles_p_long_1617)
points <- inter_ctrl$coefficients
ests <- cbind(points, confint(inter_ctrl))
plot_did_int_mip <- 
  ests %>% 
  filter(row.names(.) %in% c("postTRUE:treatTRUE", "postTRUE:treatTRUE:init_mig", "mip_mig_crimeTRUE")) %>% 
  ggplot(aes(y = row.names(.), x = points, xmin = `2.5 %`, xmax = `97.5 %`)) +
  geom_pointrange(size = 2) +
  geom_vline(xintercept = 0, col = "red", lty = 2) +
  ylab("") + xlab("Estimate") +
  scale_y_discrete(limits=rev(row.names(ests)[row.names(ests) %in% c("postTRUE:treatTRUE", "postTRUE:treatTRUE:init_mig", "mip_mig_crimeTRUE")])) +
  ggtitle("added control for mention of migrant crime")+
  xlim(-0.1, 0.4) +
  theme_xaringan()

plot_did_inter/plot_did_int_mip

```

---

# Control for MIP: Migrant Crime

```{r ind_inter_ctrl, height = 6}
points <- ind_inter_ctrl$coefficients
ests <- cbind(points, confint(ind_inter_ctrl))
plot_ind_int_mip <- 
  ests %>% 
  filter(row.names(.) %in% c("crime_share", "crime_share:init_mig", "mip_mig_crimeTRUE")) %>%
  ggplot(aes(y = row.names(.), x = points, xmin = `2.5 %`, xmax = `97.5 %`)) +
  geom_pointrange(size = 2) +
  geom_vline(xintercept = 0, col = "red", lty = 2, size = 2) +
  ylab("") + xlab("Estimate") +
  scale_y_discrete(limits=rev(c("mip_mig_crimeTRUE", "crime_share", "crime_share:init_mig"))) +
  ggtitle("added control for mention of migrant crime") +
  theme_xaringan()

plot_ind_inter/plot_ind_int_mip
```

### No: effect seems to be independent.

---

class: inverse, middle, center

## Effect on Issue Definitions

---

## Does increased exposure to crime affect likelihood to report criminal migrants as the main problem?

```{r ind_inter_dv, fig.height=6}

ind_dv_mip <- 
  feglm(mip_mig_crime ~ crime_share | wave + lfdn,  
      cluster = c("lfdn"), data = merged_data)

points <- ind_dv_mip$coefficients
ests <- cbind(points, confint(ind_dv_mip))
plot_ind_dv_mip <- 
  ests %>% 
  ggplot(aes(y = row.names(.), x = points, xmin = `2.5 %`, xmax = `97.5 %`)) +
  geom_pointrange(size = 2) +
  geom_vline(xintercept = 0, col = "red", lty = 2, size = 2) +
  ylab("") + xlab("Estimate") +
  ggtitle("Effect of exposure on reporting migrant crime as MIP") +
  theme_xaringan()

ind_inter_dv <- 
  feglm(mip_mig_crime ~ crime_share:init_mig + crime_share + init_mig | wave,  
      cluster = c("lfdn"), data = merged_data)

points <- ind_inter_dv$coefficients
ests <- cbind(points, confint(ind_inter_dv))
plot_ind_int_mip <- 
  ests %>% 
  filter(row.names(.) %in% c("crime_share", "crime_share:init_mig")) %>%
  ggplot(aes(y = row.names(.), x = points, xmin = `2.5 %`, xmax = `97.5 %`)) +
  geom_pointrange() +
  geom_vline(xintercept = 0, col = "red", lty = 2) +
  ylab("") + xlab("Estimate") +
  scale_y_discrete(limits=rev(c("crime_share", "crime_share:init_mig"))) +
  ggtitle("and interaction with prior attitudes") +
  theme_xaringan()

plot_ind_dv_mip/plot_ind_int_mip
```

### Effect larger among liberals and moderates (also huge effect).

---

class: inverse, middle, center

# Other frames

---

## All frames conditioned on initial attitude

```{r}

est_ov <- data.frame()

for (treat in 
     merged_data %>% 
     ungroup() %>%  
     select(n_articles:hungary_referendum_share, -essener_tafel_share) %>% 
     .[,apply(., FUN = mean, na.rm = T, MARGIN = 2) > 0.02] %>% # restrict to topics making up 2% or more of imm coverage on average
     colnames()){
  merged_data$treatment <- merged_data[[eval(treat)]]
  ests <- feglm(dv ~ treatment*init_mig | wave, data = merged_data, cluster = c("lfdn"))

  data.frame(
    treatment   = treat,
    treat_est   = ests$coefficients[["treatment"]],
    treat_lower = confint(ests)["treatment",][[1]],
    treat_upper = confint(ests)["treatment",][[2]],
    ia_est   = ests$coefficients[["treatment:init_mig"]],
    ia_lower = confint(ests)["treatment:init_mig",][[1]],
    ia_upper = confint(ests)["treatment:init_mig",][[2]]
  ) %>% 
  rbind(est_ov, .) -> est_ov

}

est_ov %>% 
  pivot_longer(cols = treat_est:ia_upper, names_to = c("estimate", ".value"), names_pattern = "(.*)_(.*)") %>% 
  filter(estimate == "ia") %>% 
  ggplot(aes(x = est, xmin = lower, xmax = upper, y = treatment)) +
  geom_pointrange() +
  geom_vline(xintercept = 0, col = "red", lty = 2) +
  theme_xaringan() +
  ggtitle("Interaction Effect")

```

---

class: center, inverse, middle

## Other

---

## Is effect of exposure dependent on past exposure?

```{r pret_ind, fig.height=6}

ind_pretind <- 
  feglm(dv ~ crime_share * crime_share_lag | wave,  
      cluster = c("lfdn"), data = merged_data)

points <- ind_pretind$coefficients
ests <- cbind(points, confint(ind_pretind))
plot_ind_pretint <- 
  ests %>% 
  ggplot(aes(y = row.names(.), x = points, xmin = `2.5 %`, xmax = `97.5 %`)) +
  geom_pointrange(size = 2) +
  geom_vline(xintercept = 0, col = "red", lty = 2, size = 2) +
  ylab("") + xlab("Estimate") +
  ggtitle("Interaction with past exposure") +
  scale_y_discrete(limits=rev(c("crime_share_lag", "crime_share", "crime_share:crime_share_lag")),
                   labels = rev(c("Lagged Crime Exposure", "Crime Exposure", "Crime Exposure X Lagged Crime Exposure"))) +
  theme_xaringan()

plot_ind_pretint

```

### Yes.

(Finding is sensitive to individual fixed effects.)

---

## Model Bild readers vs. other conservatives

```{r}

inter <- 
  feglm(dv ~ post * treat * init_mig, 
      cluster = c("lfdn"), 
      data = gles_p_long_1617 %>% 
        filter(`1661a_bin` | `1661c_bin` | `1661f_bin`))

points <- inter$coefficients
ests <- cbind(points, confint(inter))

plot_did_inter_cons <- 
  ests %>% 
  filter(row.names(.) %in% c("postTRUE:treatTRUE", "postTRUE:treatTRUE:init_mig")) %>% 
  mutate(name = fct_reorder(row.names(.), desc(row.names(.)))) %>% 
  ggplot(aes(y = name, x = points, xmin = `2.5 %`, xmax = `97.5 %`)) +
  geom_pointrange() +
  geom_vline(xintercept = 0, col = "red", lty = 2) +
  ylab("") + xlab("Estimate") +
  ggtitle("Control: conservative readers") +
  theme_xaringan()

plot_did_inter_cons

```

---

# Relationship of Issue Importance and Attitude

```{r}

merged_data %>% 
  filter(!is.na(`1130_clean`) & !is.na(`1140_clean`)) %>% 
  ggplot(aes(x = `1140_clean`+3, y = `1130_clean`)) +
  geom_violin(aes(x = as.factor(`1140_clean`))) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) +
  theme_xaringan()

```