### 2022-06-28 pres/Quick Modeling Summary for WIll/Quick Modeling Summary for WIll.Rmd

---
title: "Modeling Report for Will"
author: Nicolai Berk
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = FALSE, 
  message = F, 
  warning = F, 
  cache = T, 
  # out.width="80%", 
  fig.height=4, 
  fig.width=6, 
  fig.align='center', 
  eval = T, 
  dev = "svg"
  )
# knitr::opts_knit$set(root.dir = 'C:/Users/nicol/Dropbox/PhD/Papers/Bild/github/bild')

library(tidyverse)
library(dplyr)
library(data.table)
library(fixest)
library(lubridate)
library(patchwork)
library(here)

# load self-made functions
for (filename in list.files(here("code/functions"))){
  source(here(paste0('code/functions/', filename)))
}

```

# OLS

```{r did_imm_att}

DiDModels(dv_name = "1130_clean", 
          ctrl_var  = "cond",
          ctrl_name = "Initial Migration Att.")

```

---

# Tobit (no FEs, no Clustering)

```{r tobit_direct, results = 'markup'}

TobitDiD("1130_clean")

```

---

# Effect dependent on prior position

```{r}

InterLinear()

```

---

# Descriptive Plot

```{r}

DiDDist()

```

---

# Current Questions

- Interaction effect seems like an artifact?
- Main problem the inclusion of FEs similar to e.g. `fixest`.
- Seems like best estimation strategy is ordinal logit/probit.