### 2022-08-22 paper/Manuscript_Framing.Rmd
Update abstract, add change across time

---
title: "Framing Effects in the Wild"
author: "Nicolai Berk"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    toc: false
    fig_caption: yes
    includes:
      in_header: preamble.tex
# subtitle: Manuscript prepared for the 12th Annual Conference of the European Political Science Association in Prague, Czech Republic
fontsize: 12pt
geometry: margin=1.5in
abstract: "Does news framing affect public opinion? While a large body of experimental work suggests that individual attitudes can be easily altered by emphasizing different aspects of an issue, little is known about how these studies translate into the real world. I exploit a rare natural change of news framing in the major German tabloid newspaper *Bild* to study whether experimental estimates of framing effects translate into the real world. In a first step, I apply supervised transformer models to a novel dataset of 2.5M news articles and provide evidence of a substantial shift in the paper's migration framing following an editorial change. Second, I employ a difference-in-differences design using panel data from over 16,000 individuals in ten waves to assess the precise impact on migration attitudes. The findings show that, despite a major change in migration framing, immigration attitudes and related variables remain unaffected. News framing only seems to be effective within limited opportunity windows during the emergence of political issues."
bibliography: C:/Users/nicol/Dropbox/Studium/BibTex/PhD.bib
---

```{r setup, include = F}

knitr::opts_chunk$set(echo = F, message = F, warning = F, cache = T, fig.height=6, fig.width=10, fig.align='center', eval = T)

library(tidyverse)
library(dplyr)
library(data.table)
library(fixest)
library(lubridate)
library(patchwork)
library(here)

# load self-made functions
for (filename in list.files(here("code/functions"))){
  source(here(paste0('code/functions/', filename)))
}

```

WORD COUNT: `r wordcountaddin::word_count(here("paper/Manuscript_Framing.Rmd"))`

*I am thankful for extensive feedback and other support on this manuscript from Christian Arnold, Daniel Bischof, Philipp Broniecki, Jóhanna Ýr Bjarnadóttir, Elias Dinas, Florian Foos, Violetta Haas, Felix Hagemann, Julian Heid, Christina Gahn, Heike Klüver, Martin Vinæs Larsen, Matthew Levendusky, Will Lowe, Thomas M. Meyer, Rune Slothuus, Judith Spirig, Antonio Valentim as well as participants in the EPSIP Colloquium at HU Berlin, CIVICA EGN Workshop at the EUI, CompText Conference 2022, and the Political Behaviour Section at Aarhus University. All remaining confusions are my own.*

<!--# I also thank the two RAs, Robin Groß and Nathalie Aerts, for annotating the news articles.  -->

# Introduction: Does Framing Matter?

Does news framing affect public opinion? While a large body of experimental evidence suggests that the emphasis of a certain consideration can substantially alter issue attitudes, we know little about how these findings translate into real-world environments. Prior work has shown that the external validity of such experiments might be seriously complicated by environments of diverse frames [@Chong2013] and actors [@Slothuus2010], as well as the seeking and avoidance of specific content [@Lau2021a; @Taber2006]. Contemporary news environments provide consumers with especially diverse content and maximize their abilities to choose specific outlets and avoid others [@Bennett2008], thereby increasing the gap between the real-world environment and experimental studies. Additionally, the very nature of experiments is prone to over-estimate effects in the real world [@Barabas2010].

Following similar considerations, Donald Kinder called for framing research to extend its toolkit: "[E]xperimental results can always be questioned on their generalizability, and framing effects are no exception. [...] A more balanced reading of frame effects requires methodological diversification, experiments and studies oriented to the world outside" [-@Kinder2007, 157; see also @Busby2019; @Leeper2020]. To the author's knowledge, no study has yet answered this call to provide an estimate of framing effects in the real world.

I address this gap in the literature by exploiting a rare and naturally occuring shift in the migration framing of Germany's largest tabloid newspaper, *Bild*. The case is ideal to assess the everyday influence of issue frames on public opinion, as one outlet substantially changed it's framing while other papers maintained theirs. By employing BERT models for the measurement of crime frames in a corpus of 2.5 million newspaper articles, I am able to clearly identify the treatment - something that is usually left to experimental studies. The fielding of the German Longitudinal Election Study in the same period provides the rare opportunity to precisely assess the effect of this changing framing on readers' migration attitudes in a *difference-in-differences (DiD) design*.

The results indicate that even this strong change in news framing did not affect consumers' immigration attitudes. The estimated effect is statisticially indistinguishable from zero, and significantly smaller than even minimal relevant effect magnitudes. Other variables, such as related issue attitudes, issue importance, and voting intentions are similarly unaffected. Additional evidence suggests that readers do also not react by turning away from the newspaper. Overall, these results fundamentally challenge the external validity of framing effects research and suggest that the influence of individual news outlets on public opinion is rather limited. Even strong changes in news framing do not affect political attitudes or news consumption.

The paper will proceed as follows: in the next section I will briefly introduce the theoretical framework of emphasis framing and derive hypotheses about the impact of news framing on political attitudes. I will then discuss the limited external validity of framing effects studies, before I turn to the research design, including case, measurement, and estimation strategy. The results section will then showcase evidence from difference-in-differences models assessing the impact of changing migration framing on migration attitudes. Lastly, I discuss the wider implications of the findings for the study of framing effects and our understanding of the role of news media in the opinion formation process.

# Emphasis Framing

"Framing is the process by which a communication source, such as a news organisation, defines and constructs a political issue or public controversy" [@Nelson1997Rally]. A "frame" is then the operating unit of this process, defining the issue at hand in a specific way. Different forms of frames have been conceptualised. When I talk about 'frames' here, I refer to *emphasis frames*. These frames emphasise certain topics in relation to an issue, guiding the recipient to think about the issue with those considerations in mind that are promoted by the frame [@Leeper2020, 152]. For example, issues of increasing welfare contributions might be discussed with reference to inequality and providing chances to the poor, or by mentioning that higher welfare benefits might result in higher taxes. Individuals' support of a statement will differ dependent on the frame presented [@sniderman2004structure].

Work using the emphasis framing concept of framing - as opposed to other definitions like equivalence frames [@Kahneman1979prospect], episodic frames [@Iyengar2005], or generic frames [@DeVreese2005] - mostly builds on the expectancy-value model developed by @Ajzen2000. This model suggests that an attitude on a given issue is a function of two things: considerations relating to that object and their respective weights. The evaluation of the issue is subdivided into a number of considerations [a so-called "frame in thought"; @Chong2007a, 105; @Kinder1996], which will carry clear evaluations. For example, to form an opinion about whether immigration to one's country should be restricted, a person might consider the humanitarian conditions in the countries of origin (negative consideration regarding restrictive migration policy), as well as the risk of increased crime (positive consideration; note that the model does not differentiate whether a consideration is grounded in objective reality). Individuals weigh each of these considerations to form an overall opinion on the subject.

Figure \ref{fig:cogStor} visualizes this logic. A given issue is associated with an attitude $y$. The issue is related to a number of different considerations, each of which carries an associated evaluation $x_i$, e.g. that crime is bad. Each of these considerations is more or less associated with the issue. Based on the strength of this association $w_i$, a given consideration's evaluation is more or less reflected in the issue evaluation $y$, which is equal to the weighted sum of considerations.

```{=tex}
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{vis/CognitiveStorage.png}
    \caption{The cognitive evaluation of an issue $y$ is determined by the strength of association $w_i$ with other concepts with existing evaluations $x_i$.}
    \label{fig:cogStor}
\end{figure}
```
How does this model incorporate news content and other forms of political communication? News content can emphasize different considerations of an issue [so called "frames in communication" @Kinder1996]. This strengthens the cognitive association of the issue at hand with the raised consideration. As a result, this specific consideration factors more prominently in an individuals' attitude formation process: "media frames influence opinions by stressing specific values, facts, or other considerations, endowing them with greater apparent relevance to the issue than they might appear to have under an alternative frame" [@Nelson1997Rally, 569].

Applying this to the issue of migration policy and the consideration of criminal migrants, one can imagine a news consumer confronted with an increased volume of news about crimes committed by migrants. The value-expectancy model predicts that the consumer forms a cognitive definition of migration that is increasingly associated with the issue of crime, meaning this consideration features more prominently when forming an opinion about the appropriate level of immigration policy. As a result, their opinion about migration policies should become more restrictive.

***Association hypothesis*****: When an individual is exposed to a specific consideration about an issue, their association of the issue with this consideration should be strengthened.**

***Evaluation hypothesis*****: When an individual is exposed to a specific positive/negative consideration about an issue, their attitude about the overall issue should become more positive/negative.**

# Limits of Framing Experiments

Research on framing has shown for decades that citizens' attitudes and their issue associations are highly reactive to minor changes in the emphasis of specific considerations. However, it remains unclear how well these experimental findings translate into the real world. Especially when it comes to the exposure to frames in communication, translating framing effects from experimental studies is only possible under strong assumptions: While framing studies usually simply present a news article or other forms of political content to respondents, news content in the real world competes with other news and entertainment content to reach its audience. Additionally, changes in news framing itself might affect how likely readers are to consume news content: readers might be less likely to consume frames challenging their existing attitudes and reject news outlets containing such frames (and vice versa seek out attitudinally congruent information). This idea of 'selective exposure/avoidance' [@Bennett2008], or a a 'confirmation bias' [@Taber2006] brings with it a major endogeneity problem making purely correlational evidence of framing effects rather uninformative.

Even if we assume that the consumer receives the content, e.g. buys a newspaper, visits its website, or is exposed otherwise, it is not clear whether the content is actually consumed. E.g., a reader might well buy a newspaper but skip an article after reading the headline because the topic is not of interest or the frame is at odds with their existing attitude [@Lau2021a]. As most framing studies do not offer choice of content to respondents, engagement remains an assumption in experimental studies. Additionally, even when consuming the content, motivated reasoners presented with a plethora of news content about an issue will spend less time and cognitive energy on counter-attitudinal arguments than those supporting their viewpoint, lending more importance to considerations raised by amenable content [@Taber2006]. Experimental studies presenting only one frame per issue cannot take this bias into account.

These assumptions of exposure, consumption, and equal engagement in classic framing studies already represent a high hurdle for the translation of experimental estimates of framing effects into the real world. However the differences do not end there. Consumption might result in better or worse comprehension in the real world because respondents might pay more or less attention to real news content compared to experimental treatments [@Barabas2010]. Even after understanding the frame, there might be differences in how easily this translates into changing cognitive associations, given that those exposed might take real-world content more seriously.

To summarize, framing might have very different consequences in real-world environments than their experimental emulations. When we take the experimental evidence as an approximation of the effectiveness of framing in the real world, we make strong assumptions about citizens' exposure to, consumption of, and engagement with the frames in news content. Experiments can only incorporate few characteristics of the real world at a time. In order to test the external validity of framing effect studies, it is necessary to study observational data.

## Existing Observational Studies of Framing Effects

Several scholars have arrived at this conclusion and called for the assessment of framing effects beyond experimental studies [@Busby2019; @Kinder2007; @Leeper2020], yet the call has rarely been answered. While a large body of media effects research exists, it has not paid specific interest to framing, instead assessing the impact of changing news exposure broadly in observational designs [see e.g. @Foos2022; @Grossman2022 @Levendusky2022; @Spirig2020] and field experiments [@Bail2018; @Broockman2022; @Guess2021].

The precious small body of existing research on framing effects is often riddled with endogeneity concerns or compound treatments. For example, @Jerit2008 investigates the effects of framing on public opinion compared to addressing the opponent's issue. She finds that reframing is less effective than engagement with the opposite parties' frames, a finding that suggests that framing is not very effective. However, this study does not provide an estimate of a causal effect, instead restricting itself to the assessment of correlational evidence. This is especially problematic regarding framing effects, as news content can affect consumers' inclination to consume news (see preceding section), and political actors might react to public opinion. Others show with similar correlational designs that only specific frames seem to drive issue importance [@Mclaren2018], but do not assess effects on attitudes.

To my knowledge, there are only two observational studies assessing the causal effect of exposure to specific content on political attitudes. @Slothuus2010b takes advantage of a sudden shift in a political parties' framing of retirement benefits, and shows that this changed framing has substantial impact on party supporters' attitudes about the issue. However, it is unclear how these findings translate to news framing effects, as it is impossible to distinguish the effect of partisan cue-taking from the persuasive effect of the frame itself. The only study which assesses the precise impact of changing news content on political attitudes comes from @Djourelova2020. She exploits a ban of the term "illegal migrant" in the *Associated Press* and shows that this language has modest effects on migration attitudes. While the study itself reveals impressively how language can shape political attitudes, it constitutes a highly specific case and a rather minor change in the framing of migration. While news stories might lack the term "illegal migrant", they might still extensively cover crimes committed by migrants and thus associate migration with crime.

Summarized, we lack an understanding of how framing affects political opinions in the real world. While experimental evidence suggests rather strong effects, it is unclear how these findings translate to the outside world. Observational evidence on media effects is abundant but has not studied precisely what the impact of framing is. The present study therefor exploits a naturally occurring change in the news framing of a particular issue *within* a specific outlet to isolate the effects of news framing on political opinions.

# Research Design

## A Rare Shift in Migration Framing

In order to test whether changing news content - rather than a shift in the consumed news outlet - affected political attitudes, I exploit an editorial change in the largest German newspaper, *Bild*. As I will show, this editorial change resulted in a substantial shift in the papers' migration framing and is hence ideal to assess the effect of within-newspaper content shifts. The fielding of a large-scale panel survey in the same period allow to assess the impact of news content on migration attitudes.

In the summer of 2015, Germany became sanctuary for around 800,000 refugees who had fled war, violence, and famine in Syria and elsewhere. The German public debated how to deal with the newly arriving and for the first time since the post-war years, a radical right party was likely to win representation in the national parliament. Surprisingly, given its traditional populist style and right-wing takes on migration, the major tabloid newspaper *Bild* promoted an open stance towards the newly arriving in this period. This was often connected to chief editor Kai Diekmann, who hosted a refugee family in his home [@BR2015Diekmann].

In December 2016, Axel Springer, the company owning Bild, announces that Diekmann will leave the company. Less than a week after, it becomes public that prosecutors investigate Diekmann for sexual harassment. While the company claimed that this was not the reason for Diekmann's departure, the process might have been sped up as a result of the investigations\footnote{According to a former editor of the newspaper I talked to, rumors in the company at the time support this interpretation of events.} [@Spiegel2017Diekmann]. The migration coverage in *Bild* severely changed when Diekmann left the newspaper in February 2017 and Julian Reichelt took over the position as editor-in-chief. The paper increasingly started painting a picture of refugees as criminals on the tabloid's front page [@Niggemeyer2018Bild; @DLF2018Reichelt]. Headlines like "I killed 40 people and want asylum" increasingly dominated the migration coverage of the newspaper. In a speech months before he took over the position as chief editor, Reichelt outlined his motivation for such a shift in migration coverage: "I can assure you: nothing has hurt [*Bild*] economically as much as our clear, humane, empathetic stance in the refugee crisis" \footnote{Reichelt in a speech at Deutschlandfunk's conference "Formate des Politischen 2016": \url{https://vimeo.com/190347766}}. He also argued that the German media shows too little tolerance towards anti-immigration attitudes and therefor excludes those holding such views from the political discourse.

This presents a unique case to study the effects of media content on public opinion. The shift in content can be clearly identified. The change of editor itself was not discussed in the media, which means any opinion changes among readers are the result of the changing content [@Spirig2020]. Additionally, the timing of the editorial change is potentially exogenous, given the pressure from allegations against Diekmann. In summary, *this case is as close as possible to a field experiment in which a specific consideration of migration coverage is amplified in the coverage of one newspaper, but not others*.

## Measuring Frames in Communication

To measure immigration framing in the news, I collected 2.5 million newspaper articles from the websites of the most important German broadsheets \textit{Frankfurter Allgemeine Zeitung (FAZ), Spiegel Online (SPON), Süddeutsche Zeitung (SZ), Die Tageszeitung (TAZ), Die Welt} and the major German tabloid \textit{Bild} for the period 2013-2019 (for the market reach of each outlet, see table 1). Following the value-expectancy model, I identify crime frames in migration coverage in a two-step process: first, I use supervised machine learning to identify content about migration. Second, I identify content about crime *within* migration coverage by applying a similar supervised classifier. This approach closely resembles the value-expectancy model, as I first identify coverage about the *issue*, before identifying the considerations emphasized in the discussion of this issue.

To identify whether an article is about migration, i employ supervised machine learning. As migration content is rare, and training data needs to be somewhat balanced, articles cannot be selected randomly for annotation[^1]. Instead, I first construct a dictionary of terms related to migration. I use dictionary extension based on German GloVe word-embeddings\footnote{Downloadable from \url{https://deepset.ai/german-word-embeddings}.} to construct a comprehensive dictionary, and apply it to the articles. Based on the relative share of migration words in an article, I draw a stratified sample of 300 articles (100 from articles with no migration terms, 100 from the quarter with most migration terms and 100 from the remainder in between) for each newspaper for a total of 1,800 papers. Then, a coder hand-coded these articles, assessing whether their main topic is related to migration.

[^1]: If migration content made up about 3% of all articles and the sample were selected randomly, one would find around 30 migration articles per 1000 articles, which provides rather limited information about the class of interest and results in worse recall, as the best guess in highly imbalanced data is simply the most common outcome.

Using this sample, a BERT transformer model\footnote{\url{https://huggingface.co/bert-base-german-cased}} is fine-tuned on a random subset of 1,400 annotated articles (training set) with a validation set of 200 articles. After fine-tuning, the model correctly classifies 95.5% of a test set of 200 articles (F1: 0.94, recall: 0.93, precision: 0.95). This classifier is then used to annotate all 2.5 million newspaper articles. For the period from 2013 to 2019, around 90,000 (3.6%) are identified to treat the issue of migration.

```{r salience, fig.cap="Share of migration content out of all content in six different newspapers."}

MigSaliencePlot()

```

Figure \@ref(fig:salience) shows the share of migration articles in each newspaper. Clear peaks in attention are visible surrounding the refugee movements of summer 2015, as well as following the large-scale sexual assaults on new years eve in Cologne 2016[^2], and a conflict over asylum policy between the conservative sister parties CDU and CSU, which almost lead to the breakdown of the governing coalition in 2018. Newspapers seem to be event-driven and follow largely parallel trends, albeit at different levels, with the centre-left \textit{SZ} discussing migration most and the tabloid \textit{Bild} least. Attention decreases in mid-2016, but remains high compared to pre-2015 levels. Surrounding the fielding period of the survey waves (gray vertical lines), most newspapers show similar levels of attention to the issue between 2.5% and 6% of all articles, with the exception of \textit{SZ}, which moves between 5% and 12.5%.

[^2]: Germany had a large debate about refugees and crime after the new year's eve 2015/16 in Cologne. According to a report of the ministry of interior in Northrhine-Westphalia, hundreds of women were sexually assaulted and robbed by men they perceived to be of "north-african/arabic" descent (see report of the ministry of interior in Northrhine-Westphalia: \\url{[https://web.archive.org/web/20160116104020/http://m.mik.nrw.de/fileadmin/user_upload/Redakteure/Dokumente/Themen_und_Aufgaben/Schutz_und_Sicherheit/160111ssia/160111berppkoeln.pdf}](https://web.archive.org/web/20160116104020/http://m.mik.nrw.de/fileadmin/user_upload/Redakteure/Dokumente/Themen_und_Aufgaben/Schutz_und_Sicherheit/160111ssia/160111berppkoeln.pdf%7D%7D))

To identify migration coverage containing the crime consideration and thereby associating migration with crime, I first fine tune a BERT classifier for German language in a similar fashion as above, this time with coverage about crime as the outcome of interest. Using a training set of 1400 articles, a validation set of 200 and a test set of 200 articles, the classifier reaches a very good performance with an F1 score of 85.7%. I then apply this classifier to classify the approximately 90,000 articles about migration identified by the first classifier. About 7% of all migration coverage consists of articles about some form of crime\footnote{See appendix 1.2 for more information about the classifier, training process and data.}. The share of migration content devoted to news about crime within a given time frame allows me to assess how prominent the consideration of crime as a result of migration was at a specific point in time. In order to give an example of the content, table \@ref(tab:exampleArticles) displays the headlines of the five migration articles within *Bild* most likely discussing crime according to the classifier. All articles refer to crimes of migrant groups, with two referring to violence in refugee homes, two discussing police raids regarding asylum fraud, and one discussing a police report about new years eve in cologne.

```{=tex}
\begin{table}
\centering
\begin{tabular}{|p{12cm}|}
\hline
Final new years eve report of Cologne police: male groups identified \\ \hline
Police searching for Northern-african offender: asylum-raid in Dortmund! \\ \hline
Search in refugee home - mega-raid addressing asylum fraud \\ \hline
Two groups attacking each other - again bloody conflict in refugee shelter \\ \hline
Cabin fever! Violence! Mass brawls! This is behind the increase in refugee clashes \\ \hline
\end{tabular}
\caption{Headlines of five migration articles most likely to concern crime, according to BERT crime classifier. Translation by author.}
\label{tab:exampleArticles}
\end{table}
```
The left panel of figure \@ref(fig:treatment) shows the share of migration coverage devoted to crime news (solid, red line), once for Bild and once for all other newspapers (dashed, black line). As we can see, *Bild*'s attention to crime news in its migration coverage substantially increased briefly following the takeover of Julian Reichelt, while it focused less on the issue preceding the editorial change. The trends run largely parallel to each other preceding the treatment, with the major exception being the few months following the events at new year's eve in Cologne 2015/16[^3]. The right-hand panel shows difference-in-difference estimates of the crime emphasis in migration coverage of one newspaper compared to all others, before and after the editorial change in Bild. One can see that *Bild* was the only newspaper which substantially changed its emphasis of the frame, while other newspapers maintained their levels of attention.

[^3]: Note that the low levels of attention to crime in Bild's migration coverage in 2016 are partially the result of changing arrangements with a German news agency *dpa*, which provides services to quickly distribute news content similar to the Associated Press. Estimates excluding content from *dpa* indicate a change in attention to crime of similar size, as can be seen in appendix XX.

```{r treatment, fig.cap="Estimated change in attention to crime content in migration coverage, Bild vs. other newspapers. Black ticks in left panel indicate survey waves."}

TreatmentTrendPlot() + 
  TreatmentDiDPlot()

```

## Measuring Frames in Mind

In order to assess the impact of this change in framing on migration attitudes, it is necessary to use data on consumers' immigration attitudes, as well as their news consumption. The 2017 Election Panel by the German Longitudinal election study luckily contains this data. This study consists of a total of 15 waves from 2016-2020, with 10,000 - 20,000 respondents per wave [@GLES2017Panel]. Thirteen of these waves contain questions on immigration attitudes, asking respondents whether immigration of foreigners should be restricted or made easier. Responses were recorded on a seven-point Likert scale. This variable was centered on zero ($-3$ - easier, $3$ - restricted).

The panel also contains a question on newspaper consumption, which was asked in seven waves. The question provides respondents with a list of the six major daily newspapers plus an option "another daily newspaper", and asks whether they had consumed the paper in print or read articles from their website, and if so, how many days in the past week.

## Estimation

I will present evidence from a *difference-in-differences* (DiD) estimator to show the effect of the editorial change - and the related content shift - on migration attitudes:

$$
y_{iw} = \alpha + \beta_1*Post_w + \beta_2 * BildReader_i + \beta_3 * Post_w * BildReader_i + \phi_w + \rho_i + \epsilon_{it} 
$$

The dependent variable $y_{iw}$ indicates a respondent's migration attitude in a given survey wave, measured on the Likert scale described above. The conditioning variable $BildReader_i$ is a simple binary variable indicating whether respondents read *Bild* in the past week and is measured in the pre-treatment wave. $Post_w$ indicates whether an interview took place preceding or following the editorial change. The estimator of interest is $\beta_3$, which indicates the change in migration attitudes of *Bild* readers following the editorial change, controlling for pre-treatment differences ($BildReader_i$), as well as general pre-post shifts in migration attitudes ($Post_w$).

## Expected Effect Size

Despite the abundance of framing effects studies, no common standard to assess the magnitude of framing effects exists. Even the correct measurement of the magnitude is a contested question, as different control groups can be compared [see @Chong2007a, 109f for a discussion]. Additionally, framing effects differ based on the polarization and salience of the issue [@Lecheler2009a], repeated exposure [@Chong2007competitiveFraming], the sender [@Slothuus2010], political knowledge [@Chong2007b], as well as the reference of certain groups [@Nelson1996]. Generally, reported effect size around 30% of the scale are not uncommon [@Slothuus2010], sometimes even over 40% [@Chong2007a, 104], but most results lie in between 10% and 25% of the scale when comparing pro and con arguments[^4] [@Lecheler2009a; @Nelson1997Psychology; @Nelson1997Rally; @Slothuus2008; @Slothuus2010]. As experimental treatments are likely designed in order to elicit strong responses, and given biases towards null effects in the production of scientific knowledge [see e.g. the recent debate regarding nudging effects: \@Maier2022], reported effects likely display a selection bias towards strong treatments. The treatment can be considered rather strong here, as the crime consideration is associated with strongly negative evaluations and the relative increase in the prevalence of this content is large \\footnote{6 percentage point increase compared to pre-treatment share of 8.36% in Bild (over 70% increase) and 6.75% in other newspapers (88.9% increase)}. Additionally, given that readers already consumed the newspaper, the qualities of the sender should rather increase the effect size, while the high salience of the issue should decrease it. Given these considerations, I expect an effect of around 10% of the scale, or 0.7 on the seven-point scale[^5], which represents a conservative estimate compared to reported experimental effects.

[^4]: As many studies do not include a control group, this is the most comparable metric.

[^5]: As the scale runs from one to seven, it spans six points wide.

As suggested by @Laken2017, I also compare the estimated effect to equivalence bounds indicating the smallest effect size of interest (SESOI, indicated by dotted orange line in the graph), as well as a conservative estimate of the experimentally expected effect. I define the SESOI at a 2.5% change, or 0.15 on the seven-point scale, which represents a fraction of the conservative experimental estimate and sets a fairly low bar given the strong treatment. This SESOI can be used to define equivalence bounds. Any effect estimated to lie within these bounds with statistical significance (i.e. a 95%-confidence interval in-between the positive and negative effect of this size) is too small to constitute a meaningful effect, as the hypothesis that the effect reaches any size of interest can be rejected [see also @Peyton2020].

# Results

## Effect on Migration Attitudes

```{r singlemodel}

gles_p_long <- 
  fread(here('data/raw/gles/Panel/long_cleaned.csv')) %>% 
  as_tibble()

postdate <- as.Date("2017-02-01")
title <- "Mean migration attitude post-Reichelt takeover"

gles_p_long["dv"] <- gles_p_long["1130_clean"]

single_model <- 
  fixest::feglm(dv ~ post*treat | lfdn + wave_ctrl, 
                data = 
                  gles_p_long %>% 
                  filter(!is.na(dv)) %>% 
                  mutate(
                    wave_ctrl = 
                      ifelse(
                        wave %in% unique(.$wave[c(1,2)]),
                        "reference",
                        wave
                      )
                  ) %>% 
                  select(dv, post, treat, lfdn, wave_ctrl),
                cluster = c("lfdn", "wave_ctrl"))

ATT <- single_model$coefficients[["postTRUE:treatTRUE"]]
t_ATT <- single_model$coeftable["postTRUE:treatTRUE", "t value"]
p_ATT <- single_model$coeftable["postTRUE:treatTRUE", "Pr(>|t|))"]

```

Figure \@ref(fig:singledv) visualizes the effect of the change in framing on immigration attitudes. Dashed vertical lines indicate the expected effect from the experimental literature. The dotted lines indicate the equivalence boundaries. If the 95% interval is contained within these boundaries, the hypothesis that the estimate is larger than the SESOI can be rejected. The point estimate indicates a `r round(ATT, 3)`-point change among Bild readers following treatment (\$t = \$ `r round(t_ATT, 2)`, \$p = \$ `r round(p_ATT, 2)`), which is around `r round(ATT*100/0.7)`% of the expected effect of 0.7. The 95% confidence interval also contains values far smaller than the SESOI of 0.15 but clearly include the vertical solid red line indicating no change. Surprisingly, the model suggests a clear null effect.

```{r singleplot, fig.cap="Effect of change in framing on migration attitudes; difference-in-differences estimate.", fig.height=2}

DiDPlot(multi = F, scaled = "raw", 
        theoretical_effect_size = 0.7, boundary_share = 0.25)

```

## Effect on The Association of Crime and Migration Attitudes

As outlined above, framing theory suggests that increased exposure to a specific consideration about an issue leads to stronger association of the issue with this consideration. In the present case, this should mean that the correlation of attitudes on immigration with attitudes about crime should increase. The GLES Panel data contains a variable asking respondents whether "the state's capabilities to fight crime should be extended, even if that would entail more surveillance", asked in three waves overlapping with the migration attitude question. Figure \@ref(fig:corplot) shows the correlation of crime- and migration attitudes in treatment and control group, pre- and post-treatment. I added a counterfactual, applying the change in the control group to the treatment group, for easier interpretation. As we can see, the treatment group actually saw a *smaller* increase in the correlation of crime and migration attitudes than the control group, *despite weaker pre-treatment association*. Following [@Nelson1996], table \@ref(tab:cortable) regresses immigration attitudes on attitudes towards crime and investigates whether this association is generally different between the two groups\footnote{These differences are however in the opposite direction as the (substantial) differences in the descriptive correlations outlined in the plot. I am currently not sure how to interpret this.}, changes post-treatment, and whether it changes more in the treatment group using interaction terms. I find minor differences between the groups, but no significant difference when interacted with the DiD-term - respondents do not seem to associate crime more strongly with migration following the change in migration coverage in *Bild*.

```{r corplot, fig.width=8, fig.height=4, fig.cap = "Correlation of crime and migration attitudes in treatment and control group, pre- (wave 1) and post-treatment (waves 3 \\& 13)."}

MigCrimeCorPlot()

```

```{r cortable}

MigCrimeCorTable()

```

## Robustness

### Effect across time

One reason for this null effect could be that a slowly unfolding long term effect is masked by initial absence of change. To assess whether an effect develops over time, figure \@ref(fig:migwaves) assesses the effect within split-samples for each post-wave compared to the pre-treatment wave. This allows the assessment of heterogeneous treatment effects across time. The black dots with associated confidence intervals indicate the estimated effect for each wave. The horizontal dashed line indicates the expected effect based on experimental evidence (0.7), whereas the dotted horizontal line indicates the SESOI of 0.15. The vertical (red) dashed line indicates the timing of the editorial change, whereas the dotted (orange) vertical line indicates the timing of the strong increase in crime framing in June 2017, several months after the takeover by Reichelt (see figure \@ref(fig:treatment)). It s visible that, over the course of the entire year 2017, the effect is clearly smaller than the SESOI, and includes zero. However, estimates for waves 7 and 8 in late 2018 and mid-2019 - more than one and a half years after the editorial change - showcase effects statistically different from zero, and including the SESOI boundary. Note however that the effect is still tiny, with a point estimate smaller than a 0.1 on the seven-point Likert scale, as well as an upper boundary just above the SESOI at ??. The waves in late 2019 and early 2020 have point estimates at similar lower levels (??, ??), but include zero again. Overall, this suggests an effect that remains below an effect size of minimal relevance.

```{r migwaves, fig.cap="Effect of Change in Framing on Immigration Attitudes across Waves"}

# add 0.9 confidence bar

DiDByWavePlot(theoretical_effect_size = 0.7, boundary_share = 0.25)

```

### Effect on other alternative dependent variables

While immigration attitudes remain unaffected, it might well be that an increased exposure to crime framing increases individual perceptions of the importance of migration, despite not changing attitudes, or that it raises concerns about the integration rather than the immigration of migrants. Similarly, support for the radical-right *AfD* as owner of the migration issue might increased as a result of increased exposure to crime frames. Figure \@ref(fig:multidv) displays standardized effects of the change in newspaper framing on immigration attitudes (as reference) and three related dependent variables, namely integration attitudes, the likelihood to report migration as the most important problem facing the country, as well as thermometer scores for the radical-right AfD (see Appendix ?? for precise variable descriptions). Dotted lines indicate a 0.2 standard deviation change, which has been defined as the standardized equivalence boundary in existing research [@Peyton2020, 599]. All estimated effects are smaller in size than 0.05 standard deviations and associated 95%-confidence intervals contain zero. This means we can reject the hypothesis of meaningful effect size, but cannot reject that the estimate is statistically different from zero. All estimates suggest a clear null effect.

```{r multidv, fig.cap="Standardised Effect of Change in Framing on Alternative Dependent Variables"}

# add 0.9 confidence bar

DiDPlot(multi = T, scaled = T, 
        theoretical_effect_size = 0.2) # 0.2 SD suggested by Peyton 2020, 599

```
# Mechanism

## Effect on News Consumption

```{r selectionModel}

m_sel <- SelectionModel()

```

As the migration framing in Bild becomes more conservative, this could also lead readers with more liberal migration attitudes to stop reading the newspaper, as they avoid information challenging their views. Figure \@ref(fig:selection) shows the share of initial readers that report reading Bild at later stages. The three different lines describe three different groups, based on their initial migration attitude. If readers reacted by selective exposure, one would expect that liberal readers' share would drop more significantly than conservative reader's share, as they actively select out of the newspaper. As one can see, that is not the case. Liberals' and conservatives' share of readers track each other closely, with those reporting liberal migration attitudes in the first wave even reporting to read Bild at *higher* shares than conservatives in later waves. What is interesting is that those reporting neutral migration attitudes seem to be more inclined to leave the newspaper: Neutral readership declines to a level of 57%, around 5 percentage points below the share reported by liberals (62%) and 3pp below the level of conservatives (60%). Regressing individual choice to read *Bild* in later waves on initial reported migration attitudes within the subsample of initial *Bild* readers shows a `r round(m_sel$coefficients[["init_mig_groupsNeutral"]], 2)*-100`% lower chance to report Bild readership among initial neutral readers, compared to liberal readers ($p =$ `r round(summary(m_sel)$coefficients["init_mig_groupsNeutral", "Pr(>|t|)"], 3)`), whereas there is no significant or substantial difference for conservative readers. Surprisingly, *Bild* readers reporting centrist immigration attitudes select out of the newspaper at higher rates than both conservative or liberal readers, whereas there is no meaningful difference between the latter two.

```{r selectionPlot, fig.cap="Share of initial Bild readers reporting readership in later waves by initial migration attitude. Dashed (red) vertical line indicates editorial change."}

OutSelection()

```

## Openness to Change

```{r MigChange, fig.cap="Change in individual immigration attitudes 2016-2018, compared to 2014-2016. Data taken from the GLES Longterm Panel (SOURCE)."}

MigChangePlot()

MigChangeTable <- MigChangePlot(return_table = T)

```

As readers did not react by selecting out of the newspaper, the question emerges *why* readers did not react to the change in coverage. A possible explanation is that individuals informed themselves about the issue and formed an attitude once the issue emerged, in this case in 2015 during the large migratory movements in Europe. If that were the case, opinions should change more from 2014 to 2016 compared to the 2016-2018 period. Figure \@ref(fig:MigChange) shows this statistic using GLES Longterm Panel data [SOURCE]. The top panel shows the absolute change in migration attitudes from 2014-2016. That is, if a respondent held an attitude of 3 (leaning conservative) on the ten-point scale in 2014 and moved to 2 in 2016, their migration attitude would change by 1. Immigration attitudes in the GLES sample shifted on average by `r round(MigChangeTable["2014-2016",][["change_mean"]], 2)` from 2014 to 2016 (lower 95% confidence interval `r round(MigChangeTable["2014-2016",][["change_lower"]], 2)`, upper `r round(MigChangeTable["2014-2016",][["change_upper"]], 2)`), while they shifted significantly less in the period 206-2018 (`r round(MigChangeTable["2016-2018",][["change_mean"]], 2)`; [`r round(MigChangeTable["2016-2018",][["change_lower"]], 2)`, `r round(MigChangeTable["2016-2018",][["change_upper"]], 2)`]).

It becomes apparent that individual migration attitudes stabilised after 2015. This evidence is in line with the idea that migration attitudes were only responsive in 2015, when the newspaper did not follow other tabloids in their aggressively negative portrayal of refugees and their impact on society at large. When the newspaper changes its migration coverage one and a half years later, the critical window of opportunity seems to have closed, as migration attitudes had crystallized.

# Conclusion: Windows of Opportunity for News Framing Effects

The presented results are robust evidence that, despite an over 70% increase of the emphasis of crime in migration content of Germany's largest newspaper, its readers migration attitudes remained surprisingly unaffected. Effect estimates remain below minimal thresholds across time. Related dependent variables such as integration attitudes, reporting migration as the most important problem, and support for the radical-right AfD do not deviate from those of other readers. Migration attitudes are not increasingly associated with crime attitudes in this group, nor are liberal Bild readers more likely to stop reading the newspaper, nor conservative non-readers more likely to report reading it.

These findings substantially question the external validity of experimental framing studies. While it is well-established that exposure to single frames can substantially affect political attitudes in survey- and laboratory experiments, this does not seem to easily translate into real world contexts.

Several potential explanations for this difference come to mind. The most likely is certainly selective exposure. Although no evidence for selective exposure on the outlet level could be found, readers might still select content *within* the outlet - that is, they skip articles contradicting their worldview or simply do not engage with the content on a cognitive level. Alternatively, the diversity of contemporary media environments might expose individuals with so many different frames that political attitudes are ultimately stabilized. The framing of individual outlets is crowded out by the many outlets which are consumed.

While these findings are worrisome for framing scholars, they are generally good news for democracy: even when outlets try, they cannot simply manipulate their audience's political attitudes with a snap of their finger. Worries about the undermining of democracy trough the takeover of partisan news outlets might be overstated, as consumers either seem to hold a critical distance to their news outlets or consume content from so many outlets that a single one has not much influence. Descriptive data about American news diets supports this interpretation [@Guess2021a].

This study suffers from several limitations regarding the generalizabilty of the case at hand. First, the immigration issue was already highly salient for one and a half years when Reichelt took over *Bild*. This means that attitudes have possibly crystallized as the result of personal reasoning, exposure to information, or cues from one's preferred party. Second, while tabloids are often seen as manipulators of public opinion [@Foos2022], they also contain less political content and might be taken less seriously when they write about politics. Readers from other outlets might react differently. Lastly, Bild readers differ from other readers in many ways, among others they are more conservative and less knowledgeable about politics (see Appendix ??). Readers of other outlets might react differently to increased exposure to crime frames.

This study offers several takeaways for future research. First, and most importantly, the study suggests caution in the generalization of experimental results to the real world. More observational research of framing effects would help in the mapping out of scope conditions for framing effects. The present study presented one approach how this can be done. On the other hand, experimental research could try to explore external validity issues further, assessing the existence of 'crowding out' effects, as well as conditions for and types of selective exposure.

# Resources