### 2023-04-13 pres/Pres_ETH.qmd
prepare Job Talk

---
title: "<font size=12>Do Framing Effects Persist Beyond the Lab?</font>"
subtitle: "Job Market Presentation | <font color='#801414'> Public Policy Group & Immigration Policy Lab </font> | ETH Zürich"
author: "Nicolai Berk"
date: "2023-03-14"
date-format: "DD.MM.YYYY"
format: 
  revealjs:
    theme: [simple, custom.scss]
    embed-resources: true
editor: visual
bibliography: "C:/Users/nicol/Dropbox/Studium/BibTex/PhD.bib"
---

## News Framing

```{r setup}

knitr::opts_chunk$set(echo = F, cache = T, warning = F, fig.align='center', out.width="100%")

# packages
library(tidyverse)
library(here)
library(patchwork)
library(fontawesome)

# load self-made functions
for (filename in list.files(here("code/functions"))){
  source(here(paste0('code/functions/', filename)))
}

# define theme params

ipl_red <- "#801414"
ipl_gray <- "#5B5B66"
ipl_bg <- '#f0f0f0'

theme_ipl <-
  theme_minimal() +
  theme(panel.background = element_rect(fill = ipl_bg, colour = ipl_bg),
        plot.background = element_rect(fill = ipl_bg, colour = ipl_bg),
        legend.key = element_rect(fill = ipl_bg))

```

![](vis/foxbanner.png){.absolute width="500" top="250" left="0"}

![](vis/cnnborder.png){.absolute width="500" top="250" right="0"}

## Evidence on Framing {#lit}

<br>

::: incremental
<!-- -   Media effects literature mostly concerned with changes in exposure or general "slant".\ -->
<!-- <font size="5"> [@Ash2020covid; @Foos2020; @Gentzkow2010]</font> -->
-   Evidence almost [exclusively from survey experiments]{style="background-color: #801414; color: #f0f0f0;"}.
-   Unclear how these generalize.\
<font size="5"> [@Barabas2010; @Hainmueller2015] </font>
-   Real world is messy:
    -   **selective exposure** <font size="5"> [@Druckman2012] </font>
    -   **weak persistence over time** <font size="5"> [@Lecheler2011] </font>
    -   **crowding out** <font size="5"> [@Chong2013] </font>
    -   **attitude crystallization** <font size="5"> [@Bechtel2015] </font>

:::

---

<br><br><br><br>

### Do Framing Effects Generalize?

---

### Editorial Change at Germany's Largest Tabloid {.smaller}

<br>

::: incremental

-   Feb 2017: *Bild* chief editor leaves over allegations of sexual misconduct.
-   New editor: "we must talk about criminal migrants".
-   Most likely case:
    - Largest German outlet.\
    <font size="6" color="#801414">1.7M print sales and 18.5M web impressions per day.</font>
    - Seen as highly influential on public opinion.
    - Very strong frame (crime) & large change.

:::

# Treatment

## Measuring Crime Framing {#mediadata}

<br>

- Web-scrape 2.5M news articles.
- Train two classifiers (crime/migration):
  + Draw two [stratified samples](#imbalance) of 1800 articles.
  + Annotate 2x1800 articles using [app interface](#app).
  + Fine-tune [BERT](#bert) transformer models.

::: aside
[Examples](#example)
:::

## Performance

<br>

::: columns
::: {.column width="50%"}
::: {.callout-note icon="false"}
## Migration classifier

-   Accuracy: 0.96
-   **F1: 0.94**
-   Recall: 0.93
-   Precision: 0.95

**3.6%** of articles about migration.
:::
:::

::: {.column width="50%"}
::: {.callout-note icon="false"}
## Crime classifier

-   Accuracy: 0.93
-   **F1: 0.86**
-   Recall: 0.92
-   Precision: 0.8

**7%** of migration content about crime.
:::
:::
:::

## Bild's Migration Coverage

<br>

```{r trendplot}

TreatmentTrendPlot(size = 2, dpa_corrected = F, palette = c(ipl_red, ipl_gray), ) +
  theme_ipl

```

# Effect

## Estimation: Difference-in-Differences {#estimation}

::: columns
::: {.column width="30%"}
```{r didtheory}

data.frame(
  X = factor(c("pre", "pre", "post", "post", "pre", "post"),
             levels = c("pre", "post")),
  Y = c(1.1, 1.3, 1.5, 1.9, 1.3, 1.7),
  universe = factor(c(rep("counterfactual", 4), rep("factual", 2)),
                    levels = c("counterfactual", "factual")),
  group = factor(c("untreated", "treated", "untreated", 
                   "treated", "untreated", "untreated"),
                 levels = c("treated", "untreated"))
) %>% 
  ggplot(aes(x = X, y = Y, 
             lty = universe, col = group,
             group = paste0(universe, group))) +
  geom_line(size = 2) + 
  geom_point(size = 6) +
  ylab("") + xlab("") +
  theme(axis.text.y = element_blank(), 
        legend.position = "None") +
  theme_void() +
  theme(panel.background = element_rect(fill = ipl_bg, colour = ipl_bg),
        plot.background = element_rect(fill = ipl_bg, colour = ipl_bg),
        legend.position = "None") +
  scale_color_manual(values = c(ipl_red, ipl_gray))
```
:::

::: {.column width="70%"}
<br>

<font size="6" color="#801414"> $y_{w,i} = \alpha + \beta * D_{w,i}+ \phi_w + \rho_i + \epsilon_{w,i}$ </font>
:::
:::

-   Attitudinal data from [GLES Panel](#gles). \
<font size="6">(ca. 15,000 respondents in 15 waves)</font>
-   Treated: *Bild* readers in pre-treatment wave.
-   Control: *never*-Bild-readers.
-   ATE = $\Delta$ attitude Bild readers - $\Delta$ control group.
-   Parallel Trends: [pre](#sec-trends-pre), [post](#sec-trends-post)

# Results

## Effect on Main Dependent Variables {#mainresults}

<br>

```{r didest, fig.width=8, fig.height=3}

DiDPlot(col = ipl_red, multi = T, scaled = T) +
  theme_ipl

```

::: aside
Dashed lines represent lower bound of experimental estimates, dotted lines indicate Smallest Effect Size of Interest [@Lakens2017].  [More...](#sesoi)
:::

## Null Effect Explanation {#ipl}

<br>

- Sample Size: >100k observations, >15k individuals.
- Measurement:
  + Independent: [LDA](#lda) confirms increase in attention to crime in Bild.
  + Dependent: Multiple outcomes, different survey, same result.
- Spillover: possible, yet attitudes more liberal.

:::aside
See @alrababa2022
:::

## Theoretical Explanations {#why}

<br>

::: {.callout-important icon="false"}
## ❌ [Selective Exposure](#sec-explain1)
:::

::: {.callout-important icon="false"}
## ❌ [Crowding out](#sec-explain3)
:::

::: {.callout-important icon="false"}
## ❌ [Over-time effects](#sec-explain2)
:::

. . .

::: {.callout-warning icon="false"}
## ❓[Attitude Crystallisation](#sec-explain4)

Far less Attitudinal Change 2016-18 than 2014-16

:::

## Contributions

<br>

. . .

::: callout-important
## No Framing Effects in Most-Likely Case

***If*** any Effects, only During Brief **Windows of Opportunity**

:::

<!-- . . . -->

<!-- ::: callout-tip -->
<!-- ## Conditions of Media Effects: Politicization -->
<!-- ::: -->

. . .

::: callout-tip
## Conditions of Media Effects: Politicization <font size="5">[@Spirig2020]</font>

:::

. . .

::: callout-tip
## Migration Attitudes: Stability and Change <font size="5">[@Kustov2021]</font>
:::

. . .

::: callout-tip
## Measurement of Framing with Supervised Machine Learning
:::

# Thank you

## Research Agenda {.smaller visibility="uncounted"}

:::: columns
::: {.column width="75%"}

::: {.callout-ipl icon="false"}
## Media and Migration

- [<font color="black">Does Migration Coverage Always Benefit the Radical Right?</font>](https://raw.githubusercontent.com/nicolaiberk/FrameCompetition/main/pres/Presentation.pdf)
- The Effect of Radical Right Violence on Media Coverage about Migration \
(with W. Krause)
<!-- - [Does Framing Affect Voting Behaviour?](https://nicolai-berk.shinyapps.io/wzb-221214/#/title-slide) -->
<!-- - [<font color="black">How Elite Behaviour affects Individual Vaccination Decisions</font>](https://raw.githubusercontent.com/nicolaiberk/PartyVac/main/pres/Presentation.pdf) -->

:::

::: {.callout icon="false"}
## NLP & ML

- [<font color="black">Measuring Party Accommodation with Supervised Machine Learning</font>](https://raw.githubusercontent.com/nicolaiberk/bild/main/paper/Manuscript_Framing.pdf)
- Bias and Imbalance in Content Classification
<!-- - [Measuring Media Bias with Cross-Domain Deep Learning (with T. Arend)](https://raw.githubusercontent.com/nicolaiberk/nlpdl_project/main/reports/final-report-latex/final-report.pdf) -->

:::

::: {.callout-ipl icon="false"}
## Potential IPL Projects

- Does Media Coverage about Migrants affect Integration Outcomes?
- Efficient Maintenance of Classification Pipelines w. Drift Detection & Active Learning
- Government Communication about Planned Refugee Homes & Local Acceptance?
- Do Local Anti-Immigration Movements Depress Immigration at an Economic Cost?

:::
:::

::: {.column width="25%"}

::: {.right}

<br>
<br>
<br>

#### Contact

[`r fa("laptop")`](nicolaiberk.com) \| [`r fa("envelope")`](mailto:nicolai.berk@gmail) \| [`r fa("github")`](https://github.com/nicolaiberk) \| [`r fa("twitter")`](twitter.com/nicolaiberk)

:::

::: {.absolute bottom=-10 right=-30 width=250}

```{r, fig.height=5, fig.width=5, fig.align="right"}
plot(qrcode::qr_code("nicolaiberk.com"), col = c(ipl_bg, ipl_red))
```

:::
:::
::::

# Resources {visibility="uncounted"}

::: {#refs}
:::

# Appendix {visibility="uncounted"}

## Why [BERT](https://huggingface.co/docs/transformers/model_doc/bert) {#bert visibility="uncounted"}

<br>

- Pre-trained on large language corpus. <font size="5">([Details](https://www.deepset.ai/german-bert))</font> \
<font color='#801414'>➡ less training data, better performance.</font>
- Uses **contextual** embeddings. \
<font color='#801414'>➡ "fish & **chips**" vs "poker **chips**".</font>
- Outperforms humans on [SQuAD](https://rajpurkar.github.io/SQuAD-explorer/).
- SOTA in text classification.

<br>

[BACK](#mediadata)

![](vis/bertarch.png){.absolute width=400 bottom=0 right=0}

## Dealing with Imbalance {#imbalance visibility="uncounted"}

::: columns
::: {.column width="70%"}

::: callout-important
## Imbalance: Little Information in Random Samples
:::

::: callout-tip
## Informed Approach

- Generate dictionary related to outcome class.
- Extend using word embeddings to minimize bias.
- Count occurences in data.
- Draw stratified sample for subsequent training.

**Ensures overweighting of minority class.**

:::
:::

::: {.column width="30%"}

<br>

```{r, fig.height=4, fig.width=3}

counts <- 
  round(rbeta(1000, 0.2, 0.5)* 10, 0) %>%
  as_tibble() %>% 
  mutate(sample_id = cut(value, 
                         c(-Inf, 0.5, 2.5, Inf), 
                         c(1, 2, 3)),
         id = 1:n())

counts$sample_T <- rbinom(nrow(counts), 1, 0.05*(counts$value+1))

counts %>% 
  ggplot(aes(x = value)) +
  geom_histogram(aes(fill = as.character(sample_T)), bins=3) +
  theme_ipl +
  scale_fill_manual(values = c(ipl_gray, ipl_red),
                    labels = c("not sampled", "sampled")) +
  theme(legend.title = element_blank(),
        legend.position = "bottom") +
  xlab("# of dict terms in text") + ylab("Observations")

```

:::
:::

[BACK](#mediadata)

## Effective Annotation using App {#app visibility="uncounted"}

```{r, fig.cap="Screenshot of Annotation App", out.width = "\\textwidth"}

knitr::include_graphics("vis/HandCodingScreenshot.png")

```

[BACK](#mediadata)

## Five most Representative Articles {.smaller visibility="uncounted" #example}

<br>

```{r crimeExamples, results="asis"}

TenMostCrime()

```

<br>

[BACK](#mediadata)

## Prevalence DPA {visibility="uncounted"}

```{r dpa}

data.table::fread(here("data/raw/newspapers/_bild_articles.csv"), 
                  encoding = "UTF-8") %>% 
  mutate(
    dpa = str_detect(text, "(dpa)")
  ) %>% 
  mutate(month = lubridate::floor_date(date, "month")) %>% 
  mutate(
    dpa = ifelse(dpa, "DPA", "Original")
  ) %>%
  ggplot(aes(x = month, fill = dpa)) +
  geom_bar() +
  theme_minimal() +
  ggtitle("Monthly number of articles, Bild online", "By origin, 2013-2019") + 
  xlab("") + ylab("N") +
  scale_fill_manual(name = "Source",
                    values = c(ipl_red, ipl_gray)) +
  theme_ipl

```

## Attention to Immigration {.smaller visibility="uncounted"}

::: {.panel-tabset}

## incl. DPA

```{r migSalience}

MigSaliencePlot(papers = c("Bild"), 
                dpa_corrected = F) +
  geom_vline(xintercept = as.Date("2017-02-01"), col = "red", lty = 2) +
  theme_ipl

```

## excl. DPA

```{r migSalienceNODPA}

MigSaliencePlot(papers = c("Bild"), 
                dpa_corrected = T) +
  geom_vline(xintercept = as.Date("2017-02-01"), col = "red", lty = 2) +
  theme_ipl

```

:::

## Treatment excl. DPA content {.smaller visibility="uncounted"}

::: {.panel-tabset}

## incl. DPA

```{r treatment}

TreatmentTrendPlot(dpa_corrected = F, palette = c(ipl_red, ipl_gray)) +
  theme_ipl + 
  TreatmentDiDPlot(dpa_corrected = F, palette = c(ipl_red, ipl_gray)) +
  theme_ipl +
  theme(legend.position = "None")

```

## excl. DPA

```{r treatmentNODPA}

TreatmentTrendPlot(dpa_corrected = T, palette = c(ipl_red, ipl_gray)) +
  theme_ipl + 
  TreatmentDiDPlot(dpa_corrected = T, palette = c(ipl_red, ipl_gray)) +
  theme_ipl +
  theme(legend.position = "None")

```

:::

## Alternative Measurement: LDA {visibility="uncounted" #lda}

![](vis/topicChanges.png)

::: aside
[BACK](#ipl)
:::

## Dependent Variables {#gles visibility="uncounted"}

### German Longitudinal Election Study (GLES) Panel <font size="5"> [@GLES2017Panel]</font>

<!-- ::: columns -->
<!-- ::: {.column width="60%"} -->

-   [Immigration Attitudes](#secdv1)
-   [Integration Attitudes](#secdv1)
-   [Crime Attitudes](#secdv2)
-   [Support for RR](#secdv2)
-   [Most Important Problem](#secdv2)

[BACK](#estimation)

<!-- ::: -->

<!-- ::: {.column width="40%"} -->

![](vis/gles.png){.absolute height=500 bottom=0 right=0}

<!-- ::: -->
<!-- ::: -->

## Dependent Variables: Immigration {#secdv1 visibility="uncounted"}

::: columns
::: {.column width="50%"}
::: {.callout-note icon="false"}
## Immigration Attitudes

-   **Opportunities for immigration** of foreigners should be
    -   extended (-3) or
    -   restricted (3).
-   Asked in **13** waves (1 pre).
:::
:::

::: {.column width="50%"}
::: {.callout-note icon="false"}
## Integration Attitudes

-   Foreigners should...
    -   completely **adapt to German culture** (-3) or
    -   be able to live according to their own culture (3).
-   Asked in **4** waves (1 pre)
:::
:::
:::

[BACK](#gles)

## Dependent Variables: Other {#secdv2 visibility="uncounted"}

::: columns
::: {.column width="50%"}
::: {.callout-note icon="false"}
## Support for radical-right AfD

-   Attitude towards **AfD**
    -   strongly dislike (-5) to
    -   strongly like (5).
-   Asked in **14** waves (1 pre)
:::

::: {.callout-note icon="false"}
## MIP: Migration

-   Binary indicator, 1 if most important problem is migration.
:::
:::

::: {.column width="50%"}
::: {.callout-note icon="false"}
## Attitude towards crime policy

-   "state should get **more competences in fighting crime**, even if this means more surveillance"
    -   strongly disagree (1) to
    -   strongly agree (5).
-   Asked in **4** waves (1 pre)
:::
:::
:::

[BACK](#gles)

## Parallel Trends: Pre {#sec-trends-pre visibility="uncounted"}

```{r pretrend, warning = F}
ParallelTrendsPlot(color_palette = c(ipl_gray, ipl_red), multiplot = F) +
  theme_ipl
```

[BACK](#estimation)

::: aside
Data from GLES Longterm Panel [@GLES2018LangfristPanel]
:::

## Parallel Trends: Post {#sec-trends-post visibility="uncounted"}

```{r posttrend}
DescPlot(legend_position = "bottom", colorful = T, col_palette = c(ipl_red, ipl_gray)) +
  theme_ipl
```

[BACK](#estimation)

## Group Differences {.smaller visibility="uncounted"}

<br>

```{r bildtable}

GroupedTable(color_palette = c(ipl_red, ipl_bg))

```

## Expected Effect Size {#sesoi visibility="uncounted"}

<br>

-   most results in framing literature between 10% and 25% of range, up to \>40%.
-   **expected effect** to meet experimental studies at 10% of range = **0.6** on 7-point Likert or **0.34 SD**.
-   **Smallest Effect Size Of Interest** [SESOI, @Lakens2017] of 2.5% of range = **0.15** or **0.08 SD**.

::: aside
Share of scale reported due to often unreported standard deviations.
:::

[BACK](#mainresults)

## Association of Migration with Crime {#assoc visibility="uncounted"}

<br>

```{r assoc, fig.height=3.5, fig.width=8, fig.cap="Bootstrapped Correlation of Crime and Migration Attitude, per wave."}

MigCrimeCorPlotBS(cols = c(ipl_gray, ipl_red)) +
  theme_ipl

```

## Selective Exposure {#sec-explain1 visibility="uncounted"}

<br>

```{r outselect, fig.height=3.5, fig.width=8, fig.cap="Share of Bild readers indicating to have consumed Bild, by wave."}
OutSelection(color_palette = MetBrewer::met.brewer("NewKingdom", 3)[c(2, 3, 1)]) +
  theme_ipl
```

[BACK](#why)

## Over-time effects {#sec-explain2 visibility="uncounted"}

<br>

```{r overtime, fig.height=3.5, fig.width=8, fig.cap="Treatment Effect, by wave."}
DiDByWavePlot() +
  theme_ipl
```

[BACK](#why)

## Crowding Out {#sec-explain3 visibility="uncounted"}

<br>

```{r crowdout, fig.height=3.5, fig.width=8, fig.cap="Effect Conditional on Consumption of Other News Sources."}
CrowdOutEstimates(vis = T, col_scale = c(ipl_red, ipl_gray)) +
  theme_ipl +
  theme(legend.position = "none")

```

[BACK](#why)

## Attitude Crystallisation {#sec-explain4 visibility="uncounted"}

<br>

```{r migchange, warning=FALSE, fig.height=3, fig.width=6, fig.cap="Average Change in Immigration Attitude, by period"}
MigChangePlot(line_plot = T, 
              color_palette = c(ipl_red, ipl_gray), 
              distances = c(2)) +
  theme_ipl +
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank(),
    text = element_text(family = "serif"),
    legend.position = "None"
  )
```

[BACK](#why)